
# Inference Prinicipal Component Analysis {#pca}

## Description of Method and Data

### Method

In data analysis, PCA reduces the dimensions in a multivariate data set. 
It maximizes variance on uncorrelated variables, which are created at the 
time of data analysis (Jolliffe and Cadima, 2016). The first principal component
extracts the maximum  explained variance or inertia from the data table.The 
second component must be orthogonal to the first, and have the second largest variance. 
The rest of the components take on the next biggest level of variance until the 
final component retains the smallest level of variance. Factor scores are the 
new values corresponding to the observations and can be projected
on to the principal components graphically. 


```{r, include=FALSE}


graphics.off()
rm(list = ls())
gc()

# devtools::install_github("herveabdi/PTCA4CATA") # you only need this line of 
#code if you haven't downloaded the package
# install.packages("gridExtra")
#install.packages("ggplotify")
# install.packages("grid")
#install.packages("tidyverse")
suppressMessages(library(devtools))
suppressMessages(library(ExPosition, quietly = TRUE))
suppressMessages(library(corrplot))
suppressMessages(library(ggplot2))
suppressMessages(library(PTCA4CATA, quietly = TRUE))

# The following three functions allow you to arrange your plots in Rmarkdown
suppressMessages(library(gridExtra))
suppressMessages(library(ggplotify))
suppressMessages(library(grid))
suppressMessages(library(tidyverse))
suppressMessages(library(knitr))
suppressMessages(library(kableExtra))
suppressMessages(library(dplyr))
```

### Data

The data set is a survey of over 927 participants, after omitting missing 
observations with missing values, on 19 variables related to 
music. Survey questions are rated on an scale 0-5 with 0 being low and 5 being high.
For example, a 5 on Music Interest is high preference for an interest in music while
a 1 on Tempo preference indicates a liking of slow tempo. The age of the 
participants is 15 to 30. The survey assessed music 
preferences.

```{r, echo=FALSE}
music45 = read.csv("Data/music2.csv")

df <- data.frame(music45)

df2 = na.omit(df)

df3=df2[,-1]
df4=df3[,-21]
df5=df4[,-20]
df6=df2[,-22]
df7=df6[,-21]



df.music.centered <- scale(df5,
                           center = TRUE, scale = TRUE)
mt25 = as.matrix(df.music.centered)

colnames(df5) <- c("Music", "Fast", "Dance", "Folk", "Country", "Classical", 
                   "Musical", "Pop", "Rock", "Metal", "Punk", "Hiphop", 
                   "Reggae", "Swing", "RnR", "Alt", "Lat", "Techno", "Opera")

colnames(mt25) <- c("Music", "Fast", "Dance", "Folk", "Country", "Classical", 
                    "Musical", "Pop", "Rock", "Metal", "Punk", "Hiphop", 
                    "Reggae", "Swing", "RnR", "Alt", "Lat", "Techno", "Opera")
```

```{r}

head(df5) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```

## Heatmap: Correlation matrix

The correlation plot show that Punk and Dance, Rock and Roll and Rock, Opera and 
Classical, and Techno and Dance are fairly correlation to each while Metal, Punk 
and Rock form a strong cluster of correlation values. There is an argument to be 
made that a covariance matrix is best for survey data, that is a matrix that is not
normalized so that questions with a large variance take up more importance than
questions with a small variance. In other words, to equalize the questions by scaling
would take away the effect of question father from the center of gravity. 



```{r correlation plot, echo = TRUE}


corrplot(cor(mt25),        # Correlation matrix
         method = "shade", # Correlation plot method
         type = "full",    # Correlation plot style (also "upper" and "lower")
         diag = TRUE,      # If TRUE (default), adds the diagonal
         tl.col = "black", # Labels color
         bg = "white",     # Background color
         title = "",       # Main title
         col = NULL) 
```

## Analysis

The variable inf.pca contains descriptive and inference data. The eigenvalues, 
for example, are the amount of variance in the corresponding principal 
components. 


```{r InPosition, include=FALSE}
library(InPosition)

```

I decided to run a seed set to generate a consistent set of random sample to 
make my functions reproducible. 

```{r}
set.seed(91)
```

A PCA with inference was run using epPCA.inference.battery() using scaled and
centered data. Permutation testing and bootstraping were run as well. 


                                      
```{r analyze 1, message= FALSE, warning=FALSE, results = FALSE}
res_pcaInf <- epPCA.inference.battery(df5, center = TRUE, scale = "SS1",
                                      DESIGN = df4$Gender, graphs = FALSE,
                                      test.iters = 999)

```

```{r showRes, echo = FALSE, message = FALSE}
#res_pcaInf
```

```{r checkRes1, echo = FALSE, message= FALSE}
#res_pcaInf$Fixed.Data
```

```{r, echo = FALSE, message=FALSE}
#res_pcaInf$Inference.Data
```



### Testing the eigenvalues

I decided to remove the code for the histograms of the eigenvalues as it showed
a p-value of much less than 0.05. Thus, the eigenvalues were shown to be 
significant. It is a good practice to personally run a permutation test on 
eigenvalues for inference PCA, as this includes inferential statistical methods 
like bootstrap that may not have intrinsic value without a visualization 
technique like a histogram of the eigenvalues. 

### Scree Plot with Permutation Testing

The scree plot is a visualization of the eigenvalues that contain the most
explained variance along with statistical significance indicated by the color
violet and non-statistical significant indicated by the color black. The 
significance level is based on a permutation test, which is a resampling without
replacement. Thus, the null hypothesis can be evaluated to be true or false. 

Since three dimensions are significant, I will be creating two sets of factor
score plots. The Scree plot is independent of the DESIGN variable. 
The Scree Plot revealed that at least three eigenvalues are important for the 
component analysis. 

The elbow-rule states that when the screen plot goes from a step decline to a flat 
horizontal, the dimension on the steep-decline, or the first three dimensions in our case, 
are of importance.

It is important to center in PCA since the mean contains the largest amount of 
explained variance. Thus, if the data is not centered, the first dimension 
would capture the variance that the mean contributed to the data set. 

```{r}
my.scree <- PlotScree(ev = res_pcaInf$Fixed.Data$ExPosition.Data$eigs,
                      p.ev = res_pcaInf$Inference.Data$components$p.vals)

```

### Row Factor Scores

```{r factor_scores}
new_color = res_pcaInf$Fixed.Data$Plotting.Data$fi.col

new_color = recode(new_color, "#84BF30" = "blue", "#305ABF" = "deeppink")

my.fi.plot <- createFactorMap(res_pcaInf$Fixed.Data$ExPosition.Data$fi, display.labels = FALSE, # data

                            title = "Music Row Factor Scores", # title of the plot
                            axis1 = 1, axis2 = 2, # which component for x and y axes
                            pch = 19, # the shape of the dots (google `pch`)
                            cex = 2, # the size of the dots
                            text.cex = 2.5, # the size of the text
                            alpha.points = 0.1,
                            col.points = new_color, # color of the dots

                            col.labels = res_pcaInf$Fixed.Data$Plotting.Data$fi.col, # color for labels of dots
                            )
```

The observation factor scores are projected on the first two dimension with the 
DESIGN variable of gender, blue for males and pink for females. It seems at first
glance to be uniformly dispersed, but subsequent analytic tools will confirm that or 
reveal subtle gender distinctions. 


```{r, echo=FALSE}
my.fi.plot2 <- createFactorMap(res_pcaInf$Fixed.Data$ExPosition.Data$fi, display.labels = FALSE, # data

                            title = "Music Row Factor Scores", # title of the plot
                            axis1 = 2, axis2 = 3, # which component for x and y axes
                            pch = 19, # the shape of the dots (google `pch`)
                            cex = 2, # the size of the dots
                            text.cex = 2.5, # the size of the text
                            alpha.points = 0.1,
                            col.points = new_color, # color of the dots

                            col.labels = res_pcaInf$Fixed.Data$Plotting.Data$fi.col, # color for labels of dots
                            )


  
fi.labels <- createxyLabels.gen(1,2,
                             lambda = res_pcaInf$Fixed.Data$ExPosition.Data$eigs,
                             tau = round(res_pcaInf$Fixed.Data$ExPosition.Data$t),
                             axisName = "Component "
                             )
fi.labels2 <- createxyLabels.gen(2,3,
                             lambda = res_pcaInf$Fixed.Data$ExPosition.Data$eigs,
                             tau = round(res_pcaInf$Fixed.Data$ExPosition.Data$t),
                             axisName = "Component "
                             )
```

As the Scree plot indicated, the Row Factor Score plot shows 20% of the variance
on the x-axis or first dimension and 14% of the variance on the y-axis or second dimension. 

```{r}

fi.plot <- my.fi.plot$zeMap + fi.labels # you need this line to be able to save them in the end
fi.plot



```


 

```{r, echo=FALSE}
# get index for the first row of each group
grp.ind <- order(df4$Gender)[!duplicated(sort(df4$Gender))]
grp.col <- new_color[grp.ind] # get the color

grp.name <- df4$Gender[grp.ind] # get the corresponding groups
names(grp.col) <- grp.name

```

 
```{r fi_mean, echo = FALSE}
group.mean <- aggregate(res_pcaInf$Fixed.Data$ExPosition.Data$fi,
                     by = list(df4$Gender),# must be a list
                     mean)

# need to format the results from `aggregate` correctly
row.names(group.mean) <- group.mean[,1] # Use the first column as row names
fi.mean <- group.mean[,-1] # Exclude the first column

```

### Row Factor Scores with Means

Here the means of the row factor scores under the DESIGN of gender show
a clear separation along the second dimension and a subtle distinction along 
the first dimension. 

As we project the variables later on, we will be able to link gender with certain
music preferences. Thus, we can make sense of the large data table with nineteen
dimensions and thousands of data points. 

```{r fimeanplot, echo=FALSE}
fi.mean.plot <- createFactorMap(fi.mean,
                                alpha.points = 0.9,
                                col.points = grp.col[row.names(fi.mean)],
                                col.labels = grp.col[row.names(fi.mean)],
                                pch = 17,
                                cex = 4,
                                text.cex = 7)

fi.WithMean <- my.fi.plot$zeMap_background + my.fi.plot$zeMap_dots + 
  fi.mean.plot$zeMap_dots + fi.mean.plot$zeMap_text + fi.labels
fi.WithMean
```

###	Row Factor Scores with Tolerance interval

The Tolerance Interval contains a p-value of .95 indicating geometrically where
95% of the population lies in the factor score map. While there is much overlap, 
female group spreads towards the third quadrant while the male group 
spread toward the first quadrant. This will further quantify how we can interpret
future graphs with specific variables going to spefific groups of observations. 

```{r}
TIplot <- MakeToleranceIntervals(res_pcaInf$Fixed.Data$ExPosition.Data$fi,
                            design = as.factor(df4$Gender),
                            col = grp.col[row.names(fi.mean)],
                            line.size = .50, 
                            line.type = 3,
                            alpha.ellipse = .2,
                            alpha.line    = .4,
                            p.level       = .95)

# If you get some errors with this function, check the names.of.factors argument in the help.

fi.WithMeanTI <- my.fi.plot$zeMap_background + my.fi.plot$zeMap_dots + 
  fi.mean.plot$zeMap_dots + fi.mean.plot$zeMap_text + TIplot + fi.labels

fi.WithMeanTI

```


### Bootstrap interval

We can also add the bootstrap interval for the group means to see if these 
group means are significantly different. The BootCube has bootstapped means 
1,000 times for each group by 19 variables.It is generated by the user before 
making a confidence interval ellipse for PCA.

The first step is to bootstrap the group means, with 1,000 iterations. Thus, 
we are resampling with replacement one thousand times, and we can calculate
means and confidence intervals on those resampled sets of data. 

Bootstrap interval is a subset of a confidence interval, which is a statistical 
descriptor that validates if the population mean or statistic falls within say 
95% of the samples. In a bootstap inerval, the samples are the resampling data
sets with replacement. 


```{r, results='hide'}
# Depend on the size of your data, this might take a while
fi.boot <- Boot4Mean(res_pcaInf$Fixed.Data$ExPosition.Data$fi,
                     design = df4$Gender,
                     niter = 1000)
# Check what you have
#fi.boot

# What is the cube? Check the first 4 tables. You don't need to include this in 
# your output, because it's a lot of junk text.
#fi.boot$BootCube[,,1:4]
```

On the plot of the bootstrap interval, the bootstrap intervals closely surrounds 
the groups means, and they are non-overlapping. Where as the tolerance interval
geogrpahically outlines 95% of the observations, the bootstrap interval outlines
95% of the means from the resampling. Thus 95% of the group means of males are
distinct from 95% of the group means of females. 
  
```{r}
# Check other parameters you can change for this function
bootCI4mean <- MakeCIEllipses(fi.boot$BootCube[,c(1:2),], # get the first two components
                              col = grp.col[row.names(fi.mean)])

fi.WithMeanCI <- my.fi.plot$zeMap_background + bootCI4mean + 
  my.fi.plot$zeMap_dots + fi.mean.plot$zeMap_dots + 
  fi.mean.plot$zeMap_text + fi.labels

fi.WithMeanCI

```



### Loadings

Loadings can be defined several ways such as the sum of squares of 1 single component
summing to 1, the sum of squares of a single variables summing to 1, or the sum of squares of
all of the variables summing to q or the total inertia, which is the total variance. 

Metal, Punk, Rock, and Alternative are separated from the other variables
along the second dimension while Techno, Hiphop, Pop, and Dance are separated 
fromthe other variables alond the first dimension. 

Since mean of males was above the x-axis, males tend to like metal, punk, rock, 
and alternative. Females tend to like everything else, especially Latino, Musical,
Folk, Swing, Classical, and Rock and Roll. These later groups go in the positive direction
toward females and the negative direction can be projected back towards males. 
Thus, males do not like Latino, Musical, Folk, Swing, Classical, and Rock and 
Roll.

Additionally, the first dimension seems to separate Techo, Hiphop, Pop, and Dance,
which tend to be fast tempo songs from the rest of the genres which range 
from moderate to slow. 

```{r, echo=FALSE}
df5.col = res_pcaInf$Fixed.Data$Plotting.Data$fj.col

df5.col[1,1] = "brown"
df5.col[2,1] = "khaki3"
df5.col[3,1] = "khaki3"
df5.col[4,1] = "brown"
df5.col[5,1] = "brown"
df5.col[6,1] = "brown"
df5.col[7,1] = "brown"
df5.col[8,1] = "khaki3"
df5.col[9,1] = "darkorange3"
df5.col[10,1] = "darkorange3"
df5.col[11,1] = "darkorange3"
df5.col[12,1] = "khaki3"
df5.col[13,1] = "brown"
df5.col[14,1] = "brown"
df5.col[15,1] = "brown"
df5.col[16,1] = "darkorange3"
df5.col[17,1] = "brown"
df5.col[18,1] = "khaki3"
df5.col[19,1] = "brown"

cor.loading <- cor(df5, res_pcaInf$Fixed.Data$ExPosition.Data$fi)


row.names(cor.loading) <- row.names(cor.loading)


loading.plot <- createFactorMap(cor.loading,
                                constraints = list(minx = -1, miny = -1,
                                                   maxx = 1, maxy = 1),
                                col.points = "white",
                                col.labels = df5.col)
LoadingMapWithCircles <- loading.plot$zeMap + 
  addArrows(cor.loading, color = df5.col) + 
  addCircleOfCor() + xlab("Component 1") + ylab("Component 2")


LoadingMapWithCircles


```

The variance of each component is plotted for the factor scores 
for the variables. We see that the biggest amount of variance comes from those 
away from the center of gravity. Thus, Metal, Punk, Rock, Alternative, Rock and Roll, 
Opera, Swing, Musical, Latino, Pop, Dance, and Hiphop capture the largest 
amounts of variance compared to the other variables. 
  
```{r, echo=FALSE}
my.fj.plot <- createFactorMap(res_pcaInf$Fixed.Data$ExPosition.Data$fj, # data
                              title = "Music Column Factor Scores", # title of the plot
                              axis1 = 1, axis2 = 2, # which component for x and y axes
                              pch = 19, # the shape of the dots (google `pch`)
                              cex = 3, # the size of the dots
                              text.cex = 3, # the size of the text
                              col.points = df5.col, # color of the dots
                              col.labels = res_pcaInf$Fixed.Data$Plotting.Data$fj.col, # color for labels of dots
)

fj.plot <- my.fj.plot$zeMap + fi.labels # you need this line to be able to save them in the end
fj.plot
```


### Dimensions 2 and 3

Below are the Components 2 and 3 plotted based on the results of the Scree Plot.
The pink dot (females) are more dispersed away from the center of gravity and in 
third and fourth quadrant. The blue dot (males) overlap in the center of gravity
and above the x-axis. The third dimension retains 11% of the total variance of the
data set. 

```{r, echo = FALSE}

fi.plot2 <- my.fi.plot2$zeMap + fi.labels2 # you need this line to be able to save them in the end

fi.plot2
```


The males and females seem to form distinct groups. Further, it is surprising
that the female is still close to the origin. However, upon closer examination
there are several female observations on the periphery, especially in the second
and third quadrant. Here, the gender DESIGN is split along the second dimension.
The third dimension shows little group distinction along the DESIGN of gender. 

```{r,echo=FALSE}
fi.mean.plot2 <- createFactorMap(fi.mean,
                                axis1 = 2,
                                axis2  = 3,
                                alpha.points = 0.9,
                                col.points = grp.col[row.names(fi.mean)],
                                col.labels = grp.col[row.names(fi.mean)],
                                pch = 17,
                                cex = 4,
                                text.cex = 7)

fi.WithMean2 <- my.fi.plot2$zeMap_background + my.fi.plot2$zeMap_dots + 
  fi.mean.plot2$zeMap_dots + fi.mean.plot2$zeMap_text + fi.labels2
fi.WithMean2

```




```{r, echo=FALSE}
TIplot2 <- MakeToleranceIntervals(res_pcaInf$Fixed.Data$ExPosition.Data$fi,
                                  axis1 = 2,
                                  axis2 =3, 
                            design = as.factor(df4$Gender),
                            col = grp.col[row.names(fi.mean)],
                            line.size = .50, 
                            line.type = 3,
                            alpha.ellipse = .2,
                            alpha.line    = .4,
                            p.level       = .95)

# If you get some errors with this function, check the names.of.factors argument in the help.

fi.WithMeanTI2 <- my.fi.plot2$zeMap_background + my.fi.plot2$zeMap_dots + 
  fi.mean.plot2$zeMap_dots + fi.mean.plot2$zeMap_text + TIplot2 + fi.labels2

fi.WithMeanTI2
```


The tolerance interval seems to show a convex hull with the protrusion of 95% of 
female observations extending into the second, third, and fouth quadrants. 


```{r, echo = FALSE}
bootCI4mean2 <- MakeCIEllipses(fi.boot$BootCube, # get the first two components
                               axis1 =2,
                               axis2 =3,
                              col = grp.col[row.names(fi.mean)])

fi.WithMeanCI2 <- my.fi.plot2$zeMap_background + bootCI4mean2 + 
  my.fi.plot2$zeMap_dots + fi.mean.plot2$zeMap_dots + 
  fi.mean.plot2$zeMap_text + fi.labels2

fi.WithMeanCI2
```


The Confidence intervals are non-overlapping, indicating a clear separation 
between groups. 

### Loadings / Circle of Correlation


```{r, echo=FALSE}
df5.col2 = res_pcaInf$Fixed.Data$Plotting.Data$fj.col

df5.col2[1,1] = "brown"
df5.col2[2,1] = "brown"
df5.col2[3,1] = "brown"
df5.col2[4,1] = "darkorange3"
df5.col2[5,1] = "darkorange3"
df5.col2[6,1] = "darkorange3"
df5.col2[7,1] = "darkorange3"
df5.col2[8,1] = "brown"
df5.col2[9,1] = "khaki3"
df5.col2[10,1] = "khaki3"
df5.col2[11,1] = "khaki3"
df5.col2[12,1] = "brown"
df5.col2[13,1] = "brown"
df5.col2[14,1] = "brown"
df5.col2[15,1] = "brown"
df5.col2[16,1] = "khaki3"
df5.col2[17,1] = "darkorange3"
df5.col2[18,1] = "brown"
df5.col2[19,1] = "darkorange3"

cor.loading2 <- cor(df5, res_pcaInf$Fixed.Data$ExPosition.Data$fi)


row.names(cor.loading2) <- row.names(cor.loading2)


loading.plot2 <- createFactorMap(cor.loading2,
                                axis1 =2,
                                axis2 = 3,
                                constraints = list(minx = -1, miny = -1,
                                                   maxx = 1, maxy = 1),
                                col.points = "white",
                                col.labels = df5.col2)
LoadingMapWithCircles2 <- loading.plot2$zeMap + 
  addArrows(cor.loading, color = df5.col2, axis1 =2,
            axis2 =3) + 
  addCircleOfCor() + xlab("Component 2") + ylab("Component 3")


LoadingMapWithCircles2

```
Like before, the second dimension shows that males like Rock, Punk, Metal, and 
Alternative while females like everything else. The third dimension shows that 
males like the same genres of Rock, Punk, Metal, and Alternative while females
like Latino, Musical, Opera, and Classical. This is consistent to what was seen 
in the first dimension. 
 
```{r, echo=FALSE}
my.fj.plot2 <- createFactorMap(res_pcaInf$Fixed.Data$ExPosition.Data$fj, # data
                              title = "Music Column Factor Scores", # title of the plot
                              axis1 = 2, axis2 = 3, # which component for x and y axes
                              pch = 19, # the shape of the dots (google `pch`)
                              cex = 3, # the size of the dots
                              text.cex = 3, # the size of the text
                              col.points = df5.col2, # color of the dots
                              col.labels = df5.col2, # color for labels of dots
)

fj.plot2 <- my.fj.plot2$zeMap + fi.labels2 # you need this line to be able to save them in the end
fj.plot2
```

* Component 2:  Rock Concert Music vs. everything else
    Interpret: Males align with the Rock Concert Music genres. 

* Component 3:	Orchestra, Folk, and Ethic (Latino) music vs. everything else. 
    Interpret: Females align with the Orchestra, Folk, and Ethic (Latino) music
    genres. 

### Contribution barplots and Bootstrap Ratios on Variables
  
Here we are resampling the factor scores for the variables with replacement to 
generate benchmarks for how stable the data variables are in the analysis, 
specifically testing the stability of the original results. If a set of variables in the 
analysis are key to the conclusions, one can see if they are shown to also be 
important in the bootstrap ratios.

```{r}
signed.ctrJ <- res_pcaInf$Fixed.Data$ExPosition.Data$cj * sign(res_pcaInf$Fixed.Data$ExPosition.Data$fj)

# plot contributions for component 1
ctrJ.1 <- PrettyBarPlot2(signed.ctrJ[,1],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 3,
                        signifOnly = TRUE,
                         color4bar = gplots::col2hex(df5.col), # we need hex code
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ[,1]), 1.2*max(signed.ctrJ[,1])),
                         horizontal = FALSE
) + ggtitle("Contribution barplots", subtitle = 'Component 1: Variable Contributions (Signed)')
```

```{r, echo=FALSE}
# plot contributions for component 2
ctrJ.2 <- PrettyBarPlot2(signed.ctrJ[,2],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 3,
                         signifOnly = TRUE,
                         color4bar = gplots::col2hex(df5.col), # we need hex code
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ[,2]), 1.2*max(signed.ctrJ[,2])),
                         horizontal = FALSE
) + ggtitle("",subtitle = 'Component 2: Variable Contributions (Signed)')


BR <- res_pcaInf$Inference.Data$fj.boots$tests$boot.ratios
laDim = 1

# Plot the bootstrap ratios for Dimension 1
ba001.BR1 <- PrettyBarPlot2(BR[,laDim],
                            threshold = 2,
                            font.size = 3,
                            signifOnly = TRUE,
                            color4bar = gplots::col2hex(df5.col), # we need hex code
                            ylab = 'Bootstrap ratios',
                            horizontal = FALSE
                            #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
) + ggtitle("Bootstrap ratios", subtitle = paste0('Component ', laDim))

# Plot the bootstrap ratios for Dimension 2
laDim = 2
ba002.BR2 <- PrettyBarPlot2(BR[,laDim],
                            threshold = 2,
                            font.size = 3,
                            signifOnly = TRUE,
                            color4bar = gplots::col2hex(df5.col), # we need hex code
                            ylab = 'Bootstrap ratios',
                            horizontal = FALSE,
                            #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
) + ggtitle("",subtitle = paste0('Component ', laDim))

```

The the next line of code puts the figures side to side.

As stated before, males were explained to like Rock, Punk, Metal, and Alternative
as shown by the second dimension. Punk and Metal are shown to contribute to the 
second dimension importantly while Rock and Alternative do not. However, all four genres
of Rock, Punk, Metal, and Alternative have significant bootstrap ratios. Taken both
results together, Punk and Metal can be strongly argued for both contributing to the second
dimension enough and being stable enough to link them to the male gender over female gender
in terms of preference. 
  
```{r grid_ctrJ, echo = T, fig.width = 10, fig.height = 8}

grid.arrange(
  as.grob(ctrJ.1),
  as.grob(ctrJ.2),
  as.grob(ba001.BR1),
  as.grob(ba002.BR2),
  ncol = 2,nrow = 2,
  top = textGrob("Barplots for variables", gp = gpar(fontsize = 18, font = 3))
)

```


```{r, echo=FALSE}

# plot contributions for component 3
ctrJ.3 <- PrettyBarPlot2(signed.ctrJ[,3],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 3,
                        signifOnly = TRUE,
                         color4bar = gplots::col2hex(df5.col), # we need hex code
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ[,3]), 1.2*max(signed.ctrJ[,3])),
                         horizontal = FALSE
) + ggtitle("Contribution barplots", subtitle = 'Component 3: Variable Contributions (Signed)')

laDim = 3

# Plot the bootstrap ratios for Dimension 3
ba001.BR3 <- PrettyBarPlot2(BR[,laDim],
                            threshold = 2,
                            font.size = 3,
                            signifOnly = TRUE,
                            color4bar = gplots::col2hex(df5.col), # we need hex code
                            ylab = 'Bootstrap ratios',
                            horizontal = FALSE
                            #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
) + ggtitle("Bootstrap ratios", subtitle = paste0('Component ', laDim))

```

The second takeaway from above is that female survey participants showed a preference for 
Latino, Musical, Opera, and Classical as explained by the first and third dimensions. 
Opera and Classical contribute to the both the first and third dimensions while
all four genres have significant bootstrap ratios for the first dimension and Musical, Opera, 
and Classical hve significant bootstrap ratios for the third dimension. 

Taken together, we can say that Opera and Classical contribute importantly to the first and 
third dimension and have significant bootstrap ratios for those dimension. Thus, we can
strongly argue that females have a preference for Opera and Classical over male 
survey participants. 

```{r grid_ctrJ 2, echo = FALSE, fig.width = 10, fig.height = 8}

grid.arrange(
  as.grob(ctrJ.3),
  as.grob(ba001.BR3),
  ncol = 2,nrow = 2,
  top = textGrob("Barplots for variables", gp = gpar(fontsize = 18, font = 3))
)

```
### Save figures to PPT

The following code can give you a .pptx file with all the
figures saved in the directory.

*REMEMBER: Never use a screen shot*
  
```{r}
# Here we can save all figures to a PowerPoint
#savedList <- saveGraph2pptx(file2Save.pptx = 'AllFigures_inf5', 
                            #title = 'All Figures for inference5', 
                            #addGraphNames = TRUE)

```

## Summary

Component 1
Rows: 		Survey Participants		
Cols: 		Fast Songs VS Slower to moderately tempo songs
Interpret: Aside from Gender, fast/dance music differs from vs. slower to 
moderately tempo songs


Component 2
Rows: 		 Survey Participants 
Cols: 		Rock Concert Music
Interpret: 	Rock Concert Music differs from traditional and popular party music

Males tend to prefer Rock, Punk, Metal, and Alternative as told by the second 
dimension while females tend to like Latino, Musical, Opera, and Classical as 
told by the first and third dimensions. 

Bibliography
1. https://royalsocietypublishing.org/doi/10.1098/rsta.2015.0202


