
# Correspondence Analysis {#ca}


Correspondence Analysis is similar to PCA, but is meant for qualitative data.
Factor scores are generated for rows and columns, and both sets of factor 
scores have the same variance. Thus, they can be plotted on the same map for the 
analysis. CA also has distributional equivalence. Thus, merging two rows with the 
same profile does not affect the output and the output would be identical. A profile
is a relative proportion versus an absolute number as in PCA. While PCA minimizes the 
sum of square of distance, CA minimized the sum of squares of mass times distance squared
or Inertia. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, fig.width = 10)

graphics.off()
rm(list = ls())
gc()
#devtools::install_github("herveabdi/PTCA4CATA") # you only need this line of code if you haven't downloaded the package
# install.packages("gridExtra")
# install.packages("ggplotify")
# install.packages("grid")
library(devtools)
library(InPosition)
library(ggplot2)
library(corrplot)
library(PTCA4CATA)
library(data4PCCAR)
library(knitr)
library(kableExtra)
library(dplyr)

# The following three functions allow you to arrange your plots in Rmarkdown
library(gridExtra)
library(ggplotify)
library(grid)
```

The data set contains 120 consumers who evaluated their emotions upon tasting
5 types of low income sausages. The objective was to then related these
reported emotions to their sensory profiles. The participants were all
female from either Mexico City or Monterrey. 

The data table is a contingency table. Thus, variables are on the rows and 
columns while observations are the items of the table. Further, in a contingency
table, the observations are independence from one other. 

The research question of this section is: Do types of low-income sausages differ
in the emotional responses people attribute to them?


```{r data_set, echo=FALSE}
# Get the data ----

music_5 <- read.csv("data/Sausages.csv")


music_6 = data.frame(music_5)
# The active data set
X <- as.matrix(music_6)

df10 = X[,-1]

df11 = music_6[,-1]

rownames(df11) = music_6[,1]


num <- as.numeric(unlist(df11))
mt1 <- matrix(unlist(num), nrow = 5, byrow = FALSE)

rownames(mt1) = music_6[,1]
colnames(mt1) <- c("Happy", "Pleasant", "Unpleasant", "Salivating", "Famished", "Refreshed", "Desired", "Soothed", "Comforted", "Disgusted", "Energetic", "Joy", "Impressed","Interested","Irritated","Melancholic", "Nostalgic", "Relaxed", "Revital", "Romantic","Sad","Sensual",
"Espiritual","Thirsty","Well-being","Guilty")

mt2 = mt1[-5,1:26]
```

The observations for Happy emotions for the Alpino sausages is 27. Each observation 
has one sausage type and one emotion survey response. 

```{r}
head(mt1) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```

```{r, echo=FALSE}
mt2a=mt2[-4,]

mt3a = mt1[-1,1:20]

mt3b = mt3a[-1,1:20]
mt3c = mt3b[-1,1:20]
mt3d = mt3c[-1,1:20]

data.frame.3d = data.frame(mt3d)

df3e = t(data.frame.3d)

rownames(df3e) = c("Duby")

mt4=mt1[,21:26]

mt5=mt4[-5,]

df5b = data.frame(mt5)

# A supplementary points: Duby
HApunct <-df3e
# The supplementary variables
otherPunct <- df5b


```

## The data pattern
The data pattern using the chi-square is in contrast to a correlation matrix seen in PCA.
The chi-square test for independence is conducted to test for the independence of the
rows (sausage type) and columns (emotional response). It is the sum of squares for the 
observed value minus the expected value, all squared, divided by the expected value. 

The degrees of freedom are calculated by the number of rows minus one times the 
number of columns minus one of the contingency table. The number of observations 
dictates the effect of the chi-square distribution. Thus, a small number of observations
on the order of a few hundred would lead to a smaller effect than a million obsrvations. 

```{r GetChi2-1, include= FALSE, warning=FALSE}
# get Chi2 -- we can use the available package to get the Chi2
chi2    <-  chisq.test(mt1)
```

The p-value for the chi-square is non-significant. Thus, the Null hypothesis is true
that the rows are independent of the columns. In other words, there is no deviation 
from independence, partly due to the small number of observations. 
```{r}
chi2 
```

Here we are using frequencies of qualitative data while in PCA we used quantitative data. 
The chi-square statistic is divided by the total sum of the data while maintaining
the data structure of the table versus summing all the chi-square statistics to one number.
Chi-square becomes a variance between what is expected and what is observed. Just like PCA
decomposes a variance, CA decomposed chi-square. 


```{r GetChi2-2, echo = FALSE}
# Components of chi2: the chi-squares for each cell before we add them up to compute the chi2
Inertia.cells <- chi2$residuals / sqrt(sum(mt1))
# To be Plotted

# You can always compute it directly from the data
Z <-  mt1 / sum(mt1) # observed 
r <- as.matrix(rowSums(Z)) # expected for each row
c <- as.matrix(colSums(Z)) # expected for each column
# Inertia.cells 
test.Inertia.cells <- diag( as.vector(r^(-1/2)) ) %*% 
                     (Z - r%*%t(c) ) %*% diag(as.vector(c^(-1/2)))
```

Here is the heatmap for the chi-square distribution on this data set. It is a data
visualization of the chi-square values for each set of observations. It is not a correlation 
matrix but rather explains variance. The equivalent of the correlation coefficient in CA
is the Phi prime, squared value, which is a eigenvalue in the same dimension by chance. 

Bafar is capturing a positive probability frequency with being Famished while
Capistrano is capturing a positive probability frequency with Desired. Duby can 
be described closer to Joy and Impressed and away from Melancholic. Chimex is closer
in Eucledian distance to Revital while Alpino is far from Revital and Bafar from 
Romantic. 

```{r corplot, echo= FALSE}
## You can also do it without transposing it (the commented code). The plot will be the same but just in portreit.
corrplot(Inertia.cells, is.cor = FALSE)

```

## Analysis


```{r analyze, cache=TRUE, results='hide', echo=FALSE}
# run CA
resCA.sym  <- epCA(mt1, symmetric = TRUE, graphs = FALSE)
resCAinf.sym4bootJ  <- epCA.inference.battery(mt1, symmetric = TRUE, graphs = FALSE, test.iters = 25)
resCAinf.sym4bootI  <- epCA.inference.battery(t(mt1), symmetric = TRUE, graphs = FALSE, test.iters = 25)

# to run a plain CA but asymmetric
# this is using the columns as the simplex (you can also use rows by running epCA with t(X))
resCA.asym <- epCA(mt1, symmetric = FALSE, graphs = FALSE)

```

Below is the fastPerm4CA, which computes a permutation test for CA when one has 
a real contingency table. Multinomial distribution is used for the resampling. 
fastPerm4CA can be used for large tables to test for inertia and for the test 
on the eigenvalues.

```{r fastCA}
res_fast_perm <- data4PCCAR::fastPerm4CA(mt1, nIter = 100, compact = FALSE)
res_fast_boot <- data4PCCAR::fastBoot4CA(mt1)

```


### Scree Plot

Here are the results from permutation with Scree plot, with violet color 
indicating significant components.The estimated p-values were added to 
the `PlotScree` function.

As seen here, permutation testing did not capture significant values for 
dimension capturing inertia. There are 4 dimension shown as there are 5 sausage types. 
Thus, the number of dimensions is equal to the number of row or columns minus 1. 

```{r scree plot, echo=FALSE}
my.scree <- PlotScree(ev = resCA.sym$ExPosition.Data$eigs,
                      p.ev = resCAinf.sym4bootJ$Inference.Data$components$p.vals)
```



```{r toplot, echo=FALSE}
# Here are the factor scores you need
Fj.a <- resCA.asym$ExPosition.Data$fj
Fi   <- resCA.sym$ExPosition.Data$fi
Fj   <- resCA.sym$ExPosition.Data$fj

# constraints -----
# first get the constraints correct
constraints.sym  <- minmaxHelper(mat1 = Fi, mat2  = Fj)
constraints.asym <- minmaxHelper(mat1 = Fi, mat2  = Fj.a)

# Get some colors ----
color4Authors <- prettyGraphsColorSelection(n.colors = nrow(Fi))

new_color = color4Authors
new_color = recode(new_color, "mediumorchid4" = "blue", "olivedrab3" = "brown", "palegreen1" ="red", "plum" = "black", "royalblue" = "khaki3" )




#new_color2 = resCA.sym$Plotting.Data$fi.col


new_color10 = matrix(new_color)
rownames(new_color10) <- c("Alpino","Bafar","Capistrano", "Chimex", "Duby")

new_color2 =  c("blue", 
                  "brown", 
                  "red", 
                  "black",
                  "khaki3")


new_color3 = resCA.sym$Plotting.Data$fj.col

new_color3 =  c( "white", 
                  "white", 
                  "white", 
                  "white",
                  "saddlebrown",
                  "white", 
                  "white",  
                  "white",  
                  "white" , 
                  "white", 
                  "white",
                  "white",
                  "white", 
                  "white", 
                  "white",  
                  "saddlebrown", 
                  "white", 
                  "white", 
                  "white", 
                  "saddlebrown",
                  "white",
                  "saddlebrown",
                  "saddlebrown",
                  "white",
                  "white",
                 "saddlebrown"
                 )
new_color11 = matrix(new_color3)

new_color4 =  c( "palegreen3", 
                  "palegreen3", 
                  "palegreen3", 
                  "palegreen3",
                  "saddlebrown",
                  "palegreen3", 
                  "palegreen3",  
                  "palegreen3",  
                  "palegreen3" , 
                  "palegreen3", 
                  "palegreen3",
                  "palegreen3",
                  "palegreen3", 
                  "palegreen3", 
                  "palegreen3",  
                  "saddlebrown", 
                  "palegreen3", 
                  "palegreen3", 
                  "palegreen3", 
                  "saddlebrown",
                  "palegreen3",
                  "saddlebrown",
                  "saddlebrown",
                  "palegreen3",
                  "palegreen3",
                 "saddlebrown"
                 )
new_color12 = matrix(new_color4)

new_color4 = c("orange")
new_color5 = c("blue","blue" ,"blue" , "blue", "pink" , "pink")
# baseMaps ----
colnames(Fi) <- paste("Dimension ", 1:ncol(Fi))
colnames(Fj) <- paste("Dimension ", 1:ncol(Fj))
colnames(Fj.a) <- paste("Dimension ", 1:ncol(Fj.a))

```


### Plot the asymmetric factor scores

```{r asym_factor_scores, echo=FALSE}
# Your asymmetric factor scores
new_color_asym = c("blue", "brown", "red", "black", "khaki3")
asymMap  <- createFactorMapIJ(Fi,
                              Fj.a,
                            col.points.i = new_color_asym,
                            col.labels.i = new_color_asym,
                            font.face.i = 'italic',
                            alpha.labels.i = 1.2,
                            alpha.points.i = .8,
                            col.points.j = resCA.sym$Plotting.Data$fj.col,
                            col.labels.j = resCA.sym$Plotting.Data$fj.col,
                            alpha.labels.j = .9,
                            font.face.j = 'italic',
                            alpha.points.j = .8)

```

A simplex is a subspace of space with constraints. The variables are in the same
space as the rows with the barycenter at the origin. A barycenter is the middle or
center of gravity of the data. Information is inversely proportional to the frequency of 
information. Thus, the variables farthest away from other row data points explain the 
biggest variance. Romantic, Melancholic, Sensual, Famished, Guilty, and Espiritual 
capture the most variance in the data, based the simplex. The eigenvalue of .01 
for the first dimension correspond to a correlation coefficient. Thus, there is 
very low correlation value along the first dimension as well as for the next three 
dimensions. 

```{r}
# Make the simplex visible
zePoly.J <-  ggConvexHull(Fj.a, 
              percentage = 1,
              col.hull = ggplot2::alpha('darkorchid',.2),
              col.line = ggplot2::alpha('darkorchid',.7),
              line.size = .4,
              alpha.hull = .2,
              names.of.factors = "Dimension ")

# Labels
labels4CA <- createxyLabels(resCA = resCA.asym)

# Combine all elements you want to include in this plot
map.I.sup.asym <- asymMap$baseMap + zePoly.J + 
                          asymMap$I_points +
                          asymMap$J_labels + asymMap$J_points + 
                         
                          labels4CA + 
ggtitle('Asymmetric Map and Simplex') 
map.I.sup.asym
```

The first dimension separates survival emotions from higher ordered emotions. 
For example, Energetic, Famished, and Salivating are on the right (survival
emotions) while Revital, Romantic, and Desired are on the left (higher ordered
emotions). The second dimension separates positive/negative responses from 
more neutral responses. For example, Disgusted, Romantic, and Nostalgic are 
separate from Desired, Salivating, and Famished. 

### Plot the symmetric plot

```{r sym_factor_scores, echo=FALSE}
# factor scores
symMap  <- createFactorMapIJ(Fi,
                             Fj,
                            col.points.i = new_color_asym,
                            col.labels.i = new_color_asym ,
                            font.face.i = 'italic',
                            alpha.labels.i = .8,
                            alpha.points.i = .8,
                            col.points.j = new_color3,
                            col.labels.j = new_color3,
                            alpha.labels.j = .9,
                            font.face.j = 'italic',
                            alpha.points.j = .8)

```

Below is a symmetric biplot of the data. Chimex has the highest frequency probability 
for Romantic compared to the other sausage types while Duby is closest for the Guilty 
emotional response. 

#### This is a biplot:
```{r biplot_f}
# Create a symmetric map with sup and correct constraints
map.IJ.sup.sym <- symMap$baseMap + # the baseMap needs to come from your mapSup to be correct (at least for this example)
  symMap$I_labels + symMap$I_points +
  symMap$J_labels + symMap$J_points + 
  ggtitle('Symmetric Map') + 
  labels4CA

map.IJ.sup.sym
```

Below is the plots separated into two. 

```{r twoplots_f, echo=FALSE}
# plot the row factor scores with confidence intervals
map.sepI.sup.sym <- symMap$baseMap  + 
  symMap$I_labels + symMap$I_points +
 
  ggtitle('Symmetric: Row') + 
  labels4CA

# plot the columns factor scores with confidence intervals
map.sepJ.sup.sym <- symMap$baseMap +
  symMap$J_labels + symMap$J_points +

  ggtitle('Symmetric: Column') + 
  labels4CA


```


If I were a grocery clerk, I would put Capistrano and Chimez together and Bafar 
and Alpino together as the Symmetric Row plot makes this split along the first dimension. 
The Symmetric Column plot splits the Survival Emotions (Guilty and Famished) from 
the  Higher Ordered Emotions (Espiritual, Romantic, Melancholic, and Sensual) along the second
dimension. 


```{r, echo=FALSE}
grid.arrange(
    map.sepI.sup.sym, map.sepJ.sup.sym,
    ncol = 2,nrow = 1,
    top = textGrob("Factor scores", gp = gpar(fontsize = 18, font = 3))
  )

```


#### Contributions and bootstrap ratios barplots

##### Contribution barplots

For CA, we plot the contributions for both rows and columns.

```{r ctr_bars, echo=FALSE}
signed.ctrI <- resCA.sym$ExPosition.Data$ci * sign(resCA.sym$ExPosition.Data$fi)
signed.ctrJ <- resCA.sym$ExPosition.Data$cj * sign(resCA.sym$ExPosition.Data$fj)


# plot contributions of rows for component 1
ctrI.1 <- PrettyBarPlot2(signed.ctrI[,1],
                         threshold = 1 / NROW(signed.ctrI),
                         font.size = 3,
                         signifOnly = TRUE,
                         color4bar = gplots::col2hex(new_color10), # we need hex code
                         
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrI), 1.2*max(signed.ctrI))
) + ggtitle("Component 1", subtitle = 'rows')

# plot contributions of columns for component 1
ctrJ.1 <- PrettyBarPlot2(signed.ctrJ[,1],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 3,
                          signifOnly = TRUE,
                         color4bar = gplots::col2hex(new_color12), # we need hex code
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
) + ggtitle("", subtitle = 'columns')

# plot contributions of rows for component 2
ctrI.2 <- PrettyBarPlot2(signed.ctrI[,2],
                         threshold = 1 / NROW(signed.ctrI),
                         font.size = 3,
                          signifOnly = TRUE,
                         color4bar = gplots::col2hex(new_color10), # we need hex code
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrI), 1.2*max(signed.ctrI))
) + ggtitle("Component 2", subtitle = 'rows')

# plot contributions of columns for component 2
ctrJ.2 <- PrettyBarPlot2(signed.ctrJ[,2],
                         threshold = 1 / NROW(signed.ctrJ),
                         font.size = 3,
                          signifOnly = TRUE,
                         color4bar = gplots::col2hex(new_color12), # we need hex code
                         ylab = 'Contributions',
                         ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
) + ggtitle("", subtitle = 'columns')

```

As referenced in the analysis above, Bafar and Chimex makes a important contribution 
to the inertia of the chi-square distribution. Famished, Melancholic, and Sensual
make an important contribution for the first dimension while. Famished, Melancholic,
Romantic, and Guilty make up an important contribution for the second dimension. 

```{r, echo=FALSE}
grid.arrange(
    as.grob(ctrI.1),as.grob(ctrJ.1),as.grob(ctrI.2),as.grob(ctrJ.2),
    ncol = 2,nrow = 2,
    top = textGrob("Contributions", gp = gpar(fontsize = 18, font = 3))
  )
Ctr.IJ <- recordPlot() # you need this line to be able to save them in the end
```

Energetic, Famished, and Salivating as well as Romantic and Desired are among 
the variable the contribution importantly to the first dimension. Disgusted and 
salivating are variable discussed above that also contributed importantly to the
second dimension. 

##### Bootstrap ratios


``` {r BR, echo=FALSE}
BR.I <- resCAinf.sym4bootI$Inference.Data$fj.boots$tests$boot.ratios
BR.J <- resCAinf.sym4bootJ$Inference.Data$fj.boots$tests$boot.ratios

laDim = 1

# Plot the bootstrap ratios for Dimension 1
ba001.BR1.I <- PrettyBarPlot2(BR.I[,laDim],
                  threshold = 2,
                  font.size = 3,
                  signifOnly = TRUE,
                  color4bar = gplots::col2hex(new_color10), # we need hex code
                  ylab = 'Bootstrap ratios'

) + ggtitle(paste0('Component ', laDim), subtitle = 'rows')

ba002.BR1.J <- PrettyBarPlot2(BR.J[,laDim],
                  threshold = 2,
                  font.size = 3,
                  signifOnly = TRUE,
                  color4bar = gplots::col2hex(new_color12), # we need hex code
                  ylab = 'Bootstrap ratios'
            
              
) + ggtitle("", subtitle = 'columns')

# Plot the bootstrap ratios for Dimension 2
laDim = 2
ba003.BR2.I <- PrettyBarPlot2(BR.I[,laDim],
                  threshold = 2,
                  font.size = 3,
                  signifOnly = TRUE,
                  color4bar = gplots::col2hex(new_color10), # we need hex code
                  ylab = 'Bootstrap ratios',
                  ylim = c(-1,3)
             
              
) + ggtitle(paste0('Component ', laDim), subtitle = 'rows')

ba004.BR2.J <- PrettyBarPlot2(BR.J[,laDim],
                  threshold = 2,
                  font.size = 3,
                  signifOnly = TRUE,
                  color4bar = gplots::col2hex(new_color12), # we need hex code
                  ylab = 'Bootstrap ratios',
                  ylim = c(-5,5)

) + ggtitle("", subtitle = 'columns')
```

(The next set of code is used to put two figures side to side).

Bafar and Chimex as stable row variables in the analysis while 
Sensual and Famished are stables column variables. 

```{r grid_BR, echo = T}
grid.arrange(
    as.grob(ba001.BR1.I),as.grob(ba002.BR1.J),as.grob(ba003.BR2.I),as.grob(ba004.BR2.J),
    ncol = 2,nrow = 2,
    top = textGrob("Bootstrap ratios", gp = gpar(fontsize = 18, font = 3))
  )

```

Here are the contribution and bootstrap ratio plots side by side.

```{r}
grid.arrange(
     as.grob(ctrI.1),as.grob(ctrJ.1),as.grob(ctrI.2),as.grob(ctrJ.2),as.grob(ba001.BR1.I),as.grob(ba002.BR1.J),as.grob(ba003.BR2.I),as.grob(ba004.BR2.J),
    ncol = 4,nrow = 2,
    top = textGrob("Contribution   &   Bootstrap ratios", gp = gpar(fontsize = 18, font = 3))
  )

```



## Summary

First and foremost, I prefer the symmetric plot do to better visualization. 

When we interpret the factor scores and loadings together, the CA revealed:

*Chimex has the highest frequency probability for Romantic compared to the 
other sausage types 

*Duby is closest for the Guilty emotional response. 

When we interpret the factor scores and loadings separately, the CA revealed:

* Component 1: Capistrano and Chimez  vs. Bafar and Alpino

* Component 2: Survival Emotions (Guilty and Famished) vs. 
  Higher Ordered Emotions (Espiritual, Romantic, Melancholic, and Sensual)
