
# Multiple Correspondence Analysis {#mca}

This method is an extension of PCA but analyzes categorical data instead. 
In some respects it is analogous to CA but uses binary data instead. MCA can 
also be adapted to quantitative data that is scored, for example, -5 to 5 and 
then binned to represent a pattern of 0 and 1 data entries. 


```{r, include = FALSE, echo=FALSE}
graphics.off()
rm(list = ls())
gc()

# Install packages ----
#install.packages("devtools")
#devtools::install_github('HerveAbdi/PTCA4CATA')
#devtools::install_github('HerveAbdi/DistatisR')
#devtools::install_github('HerveAbdi/data4PCCAR')
#devtools::install_github('HerveAbdi/R4SPISE2018') # of course!
#install.packages('prettyGraphs')
#install.packages('Matrix')
#install.packages('dplyr')
#install.packages('gridExtra')
#install.packages('grid')
#install.packages('gtable')
#install.packages('stringi')
#install.packages('printr')
#install.packages('kableExtra')
#  load the libraries that we will need
suppressMessages(library(Matrix))
suppressMessages(library(prettyGraphs))
suppressMessages(library(ExPosition))
suppressMessages(library(InPosition))
suppressMessages(library(DistatisR))
suppressMessages(library(dplyr))
suppressMessages(library(gridExtra))    # to save a table as a graph
suppressMessages(library(grid))         # that will be saved in the
suppressMessages(library(gtable))       # powerpoint with the figures
suppressMessages(library(printr))     # To pretty print tables 
suppressMessages(library(kableExtra)) # To pretty print tables
# If not installed here it is
#remotes::install_github('HerveAbdi/DistatisR')
suppressMessages(library(DistatisR))
suppressMessages(library(PTCA4CATA))
suppressMessages(library(data4PCCAR))


# Load data
music = read.csv("data/music5.csv")
# How to put
data_2=music[,-1]
rownames(data_2)=music[,1]
# How to change to a data frame and remove rows with missing variables

data_3 = data.frame(data_2)
data_4 = na.omit(data_3)

# Transform the data into factors
rawData <- apply(data_3,2,as.factor)

rawData <- as.data.frame(rawData, stringsAsFactors = TRUE)

```

## Cleaning the data


```{r, echo=FALSE}
rawData2 = data.frame(data_4)

rawData2$Age <- as.numeric(as.character(rawData2$Age))
colnames(rawData2) <- c('supVar','Music','Fast','Dance', 'Folk','Country','Classical','Musical',
                        'Pop','Rock','Metal','Punk','Hiphop','Reg','Swing','RnR','Alt',
                        'Lat','TT','Opera','Age','Gender','City','Ed')

# How to prepare data for analysis with histogram

rawData2[,2:21] <- lapply(rawData2[,2:21], as.numeric)
```

This is the same data from PCA, containing Music preferences. I cleaned the 
data by binned the data in to roughly equal groups. For example, for the question 
of Music preferences, I binned the first 4 scores into one group and left the 
score of 5 in its own group. There are 918 survey participants and 20 variables, 
including Gender. 


```{r}

hist.Mu <- hist(rawData2[,2], breaks = 20) # 1 to 4 and 5

```


```{r, echo=FALSE}
# how to equalize the data through visualization with a histogram

#hist.Mu <- hist(rawData2[,2], breaks = 20) # 1 to 4 and 5
#hist.Fa <- hist(rawData2[,3], breaks = 20) # 1 to 2, 3, and 4 to 5
#hist.D <- hist(rawData2[,4], breaks = 20) # 1 to 2, 3, and 4 to 5
#hist.Fo <- hist(rawData2[,5], breaks = 20) # 1, 2, and 3 to 5
#hist.Co <- hist(rawData2[,6], breaks = 20) # 1, 2, and 3 to 5
#hist.Cl <- hist(rawData2[,7], breaks = 20) # 1, 2, 3, 4, 5
#hist.Ml <- hist(rawData2[,8], breaks = 20) # 1, 2, 3, 4 to 5
#hist.Po <- hist(rawData2[,9], breaks = 20) # 1 to 2, 3, 4, 5
#hist.Ro <- hist(rawData2[,10], breaks = 20) # 1 to 3, 4, 5
#hist.Me <- hist(rawData2[,11], breaks = 20) # 1, 2, 3 to 5
#hist.Pu <- hist(rawData2[,12], breaks = 20) # 1, 2, 3, 4 to 5
#hist.H <- hist(rawData2[,13], breaks = 20) # 1, 2, 3, 4, 5
#hist.Re <- hist(rawData2[,14], breaks = 20) # 1, 2, 3, 4 to 5
#hist.Sw <- hist(rawData2[,15], breaks = 20) # 1, 2, 3, 4 to 5
#hist.RnR <- hist(rawData2[,16], breaks = 20) # 1 to 2, 3, 4, 5
#hist.A <- hist(rawData2[,17], breaks = 20) # 1, 2, 3, 4, 5
#hist.L <- hist(rawData2[,18], breaks = 20) # 1, 2, 3, 4, 5
#hist.T <- hist(rawData2[,19], breaks = 20) # 1, 2, 3 to 5
#hist.O <- hist(rawData2[,20], breaks = 20) # 1, 2, 3 to 5
#hist.A <- hist(rawData2[,21], breaks = 20) #<20 Y, >20 O
```

## Recoding the data
```{r, echo=FALSE}
## ----recode Age etc. -------------------------------------
rawData2[,'Music'] <- plyr::mapvalues(rawData2[,'Music'], 
                                   from = c("1", "2","3","4", "5"), to = c("1","1","1","1", "2"))
rawData2[,'Fast'] <- plyr::mapvalues(rawData2[,'Fast'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","1","2","3", "3"))
rawData2[,'Dance'] <- plyr::mapvalues(rawData2[,'Dance'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","1","2","3", "3"))
rawData2[,'Folk'] <- plyr::mapvalues(rawData2[,'Folk'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","3", "3"))
rawData2[,'Country'] <- plyr::mapvalues(rawData2[,'Country'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","3", "3"))
rawData2[,'Classical'] <- plyr::mapvalues(rawData2[,'Classical'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "5"))
rawData2[,'Musical'] <- plyr::mapvalues(rawData2[,'Musical'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "4"))
rawData2[,'Pop'] <- plyr::mapvalues(rawData2[,'Pop'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","1","2","3", "4"))
rawData2[,'Rock'] <- plyr::mapvalues(rawData2[,'Rock'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","1","1","2", "3"))
rawData2[,'Metal'] <- plyr::mapvalues(rawData2[,'Metal'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","3", "3"))
rawData2[,'Punk'] <- plyr::mapvalues(rawData2[,'Punk'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "4"))
rawData2[,'Hiphop'] <- plyr::mapvalues(rawData2[,'Hiphop'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "5"))
rawData2[,'Reg'] <- plyr::mapvalues(rawData2[,'Reg'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "4"))
rawData2[,'Swing'] <- plyr::mapvalues(rawData2[,'Swing'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "4"))
rawData2[,'RnR'] <- plyr::mapvalues(rawData2[,'RnR'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","1","2","3", "4"))
rawData2[,'Alt'] <- plyr::mapvalues(rawData2[,'Alt'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "5"))
rawData2[,'Lat'] <- plyr::mapvalues(rawData2[,'Lat'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "5"))
rawData2[,'TT'] <- plyr::mapvalues(rawData2[,'TT'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","3", "3"))
rawData2[,'Opera'] <- plyr::mapvalues(rawData2[,'Opera'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","3", "3"))
rawData2[,'Age'] <- plyr::mapvalues(rawData2[,'Age'], 
                                      from = c("15","16","17","18","19","20","21","22","23","24","25","26","27","28","29","30"), to = c("Y","Y","Y","Y","Y","Y","Y","Y", "O","O","O","O","O","O","O","O"))
rawData2[,'Gender'] <- plyr::mapvalues(rawData2[,'Gender'], 
                                      from = c("male","female"), to = c("M","F"))
rawData2[,'City'] <- plyr::mapvalues(rawData2[,'City'], 
                                      from = c("village","city"), to = c("V","C"))

rawData2[,'Ed'] <- plyr::mapvalues(rawData2[,'Ed'], 
                                      from = c("primary school","secondary school","college/bachelor degree","masters degree","doctorate degree"), to = c("1","2","3","4", "5"))



rawData3 = data.frame(rawData2)
rawData4 = rawData3[,-1]
```


```{r}
head(rawData4) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```

```{r, echo=FALSE}
## ---- cleanData-------------------------------------------
cleanData.tmp <- rawData4

cleanData.tmp <- cleanData.tmp[complete.cases(cleanData.tmp),]

cleanData.allVar <- cleanData.tmp

cleanData.varSup <- cleanData.allVar[,20:23]
cleanData        <- cleanData.allVar[,1:19]


```

## Analysis

```{r, message = FALSE}
## ----runMCA-----------------------------------------------
resMCA <- epMCA(cleanData, graphs = FALSE)
```

```{r, echo=FALSE}

## ----runMCA.sup-------------------------------------------
# recode the factors as set of 0/1 variables
testclean <- makeNominalData(cleanData)

# barycentric code for nation
#clean.Sup[,(colnames(testclean) %in% 'nation.F')] <- .5
#clean.Sup[,(colnames(testclean) %in% 'nation.V')] <- .5
#

```

```{r, echo=FALSE}
## ----runMCA.varsup----------------------------------------
#
resMCA.varSup <- supplementaryCols(
              SUP.DATA = makeNominalData(cleanData.varSup),
              res = resMCA)
colnames(resMCA.varSup$fjj) <- paste0('Dimension ',
                                1:ncol(resMCA.varSup$fjj))
```


The Scree plot reveals that three dimensions are important based on the permutation testing 
and the Kaiser line. 

```{r, echo=FALSE}
## ----Inferences --------
resMCA.inf <- epMCA.inference.battery(cleanData, 
                                      graphs = FALSE)

## ----screeMCA.inf  -------------------
scree.mca <- PlotScree(ev = resMCA$ExPosition.Data$eigs, 
        p.ev = resMCA.inf$Inference.Data$components$p.vals, 
        plotKaiser = TRUE,
        title = "MCA. Explained Variance per Dimension")

```

The heat map is done on a Burt Table, which is a frequency table that is dummy 
coded for the groups of variables. It it based on disjunctive coding, a type of 
group coding, where either a variable is coded as a 1 or 0. Rock and Metal have 
a high positive correlation value as does Rock and Punk as well as Metal and Punk. Dance
and Dance, Dance and Techno, and Rock and Rock and Roll have high levels of 
possitive correlation. In addition, Musical and Opera as well as Classical and 
Opera have notable positive correlation values. 

```{r, echo=FALSE}
## ----colors------------------------------------------
cJ <- resMCA$ExPosition.Data$cj
color4Var <- prettyGraphs::prettyGraphsColorSelection(
                                          ncol(cleanData))

## ----phi2 ---------------
# Pseudo Heat Map. Correlation ----
# We need correlation to compare with PCA
corrMatBurt.list <- phi2Mat4BurtTable(cleanData)
col <- colorRampPalette(
  c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corr4MCA.r <- corrplot::corrplot(
         as.matrix(corrMatBurt.list$phi2.mat^(1/2)), 
         method="color", col=col(200),  
         type="upper", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col = color4Var, 
         tl.cex = .9,
         tl.srt = 45, #Text label color and rotation
         number.cex = .5,
         diag = TRUE # needed to have the color of variables correct
         )
```

```{r, echo=FALSE, warning=FALSE,message=FALSE, results=FALSE}
## ----ctrVar --------
varCtr <- data4PCCAR::ctr4Variables(cJ) 
rownames(color4Var) <- rownames(varCtr)

## ----ctrVar.Tab ---------
nFact <- min(5, ncol(cJ) - 1)
#knitr::kable(round( varCtr[,1:nFact]*1000 ) )
# save table as a graph
ctrTable <- tableGrob(round(varCtr[,1:nFact]*1000))
h <- grobHeight(ctrTable)
w <- grobWidth(ctrTable)
title <- textGrob("Variable Contributions",
                 y = unit(0.5,"npc") + 0.92*h, 
                 # fine tune the position of the title 
                  just = "centre",
                  gp = gpar(fontsize = 14))
TableWithTitle <- gTree(children = gList(ctrTable, title))

```

```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ---- grid.draw ----
# Note: Potential problems with grid.draw(). If it does not plot
# recordPlot() will fail and the graph will not be saved in the powerpoint
# and will generate a strange error message
#grid.draw(TableWithTitle)
#dev.new()

```

```{r, include = FALSE, echo=FALSE}
## ----printr ------------
# As an alternative we print the contributions with a combination
#of `kable` and `printr` as:
laTable <- round(varCtr[,1:nFact]*1000)
# knitr::kable(round(varCtr[,1:nFact]*1000), caption = 'Variable Contributions')
#    %>%
#   kable_styling(latex_options = c("striped", "hold_position"), full_width = F) %>%
#  add_header_above(c(" ", "Dimensions" = nFact))
```

As referenced in the previous comments, Rock, Metal, Punk, Rock and Roll, 
Opera, and Classical are important contributors to the first dimension. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----color4Levels ------
col4Levels <- data4PCCAR::coloringLevels(
           rownames(resMCA$ExPosition.Data$fj), color4Var)
col4Labels <- col4Levels$color4Levels

## ----ctrV1-----------------------
varCtr1 <- varCtr[,1]
names(varCtr1) <- rownames(varCtr)
a0005.Var.ctr1  <- PrettyBarPlot2(varCtr1,
              main = 'Variable Contributions: Dimension 1',
              ylim = c(-.05, 1.2*max(varCtr1)),
              font.size = 5,
              signifOnly = TRUE,
              threshold = 1 / nrow(varCtr),
                                color4bar = gplots::col2hex(color4Var)
)
print(a0005.Var.ctr1)
```

Rock, Pop, Dance, and Tech, all of which were reference before, are important 
contributors to the second dimension. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----ctrV2 -------
varCtr2 <- varCtr[,2]
names(varCtr2) <- rownames(varCtr)
a0006.Var.ctr2  <- PrettyBarPlot2(varCtr2,
                    main = 'Variable Contributions: Dimension 2',
                    ylim = c(-.05, 1.2*max(varCtr2)),
                    threshold = 1 / nrow(varCtr),
                    font.size = 5,
                    signifOnly = TRUE,
                    color4bar = gplots::col2hex(color4Var)
)
print(a0006.Var.ctr2)
```


Rock, Metal, Punk, Musical, Classical, and Opera are important contributors to the 
third dimension. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----ctrV3 -------------
varCtr3 <- varCtr[,3]
names(varCtr3) <- rownames(varCtr)
a0006.Var.ctr3  <- PrettyBarPlot2(varCtr3,
             main = 'Variable Contributions: Dimension 3',
             ylim = c(-.05, 1.2*max(varCtr2)),
             threshold = 1 / nrow(varCtr),
             font.size = 5,
             signifOnly = TRUE,
             color4bar = gplots::col2hex(color4Var)
)
print(a0006.Var.ctr3)
```


To read the next graph, the data point near the axis are importantly mostly 
to that corresponding dimension. Thus, Dance, Techno, and Pop are near the 
y-axis so they all contribute heavly to the second dimension. Rock, Opera, and 
Classical are near the x-axis so they all contribute heavly to the first dimension. 
The farther away from the center of gravity indicates higher variance. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----ctrV12----------------------------------------
ctrV12 <- PTCA4CATA::createFactorMap(X =  varCtr, 
                        title = "Variable Contributions", 
                        col.points = color4Var,
                        col.labels = color4Var,
                        alpha.points = 0.5,
                        cex = 2.5, 
                        alpha.labels = 1, 
                        text.cex = 4,
                        font.face = "plain", 
                        font.family = "sans")

ctr.labels <- createxyLabels.gen(
  1,2, lambda = resMCA$ExPosition.Data$eigs,
  tau = resMCA$ExPosition.Data$t
)
a0007.Var.ctr12  <- ctrV12$zeMap  + ctr.labels
#
print(a0007.Var.ctr12)

```

In addition to the variables spotted on the heat map, the next graph captures the 
important variables based on the contribution level by setting a cut of of 1 divided
by the number of dimension (19) in out case, leaving a benchmark of 0.05263. 

In addition to what was stated before, dimension 1 and 2 include Latino, Swing, and 
Alternative. Similarly, dimension 2 and 3 capture Hiphop, Latino, and Alternative. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----getCtr12------------------------------
absCtrVar <- as.matrix(varCtr) %*% 
                    diag(resMCA$ExPosition.Data$eigs)
varCtr12  <- (absCtrVar[,1] + absCtrVar[,2]) / 
                (resMCA$ExPosition.Data$eigs[1] + 
                resMCA$ExPosition.Data$eigs[2])
importantVar <- (varCtr12 >=  1 / length(varCtr12))
col4ImportantVar <- color4Var
col4NS <- 'gray90' 
col4ImportantVar[!importantVar] <- col4NS

## ----ctrV12.ns-----------------------------------------------------------
ctrV12.imp <- PTCA4CATA::createFactorMap(X =  varCtr, 
             title = "Important Variables: Contributions", 
             col.points = col4ImportantVar,
             col.labels = col4ImportantVar,
             alpha.points = 0.5,
                        cex = 2.5, 
                        alpha.labels = 1, 
                        text.cex = 4,
                        font.face = "plain", 
                        font.family = "sans")
a0008.Var.ctr12.imp  <- ctrV12.imp$zeMap  + ctr.labels
#
print(a0008.Var.ctr12.imp)


## ----getCtr23------------------------------------------------------------
#absCtrVar <- as.matrix(varCtr) %*% diag(resMCA$ExPosition.Data$eigs)
varCtr23  <- (absCtrVar[,3] + absCtrVar[,2]) / 
   (resMCA$ExPosition.Data$eigs[3] + resMCA$ExPosition.Data$eigs[2])
importantVar23 <- (varCtr23 >=  1 / length(varCtr23))
col4ImportantVar23 <- color4Var
col4NS <- 'gray90' 
col4ImportantVar23[!importantVar23] <- col4NS

## ----ctrV23.ns-----------------------------------------------------------
ctrV23.imp <- PTCA4CATA::createFactorMap(X =  varCtr,
                                         axis1 = 3, axis2 = 2,
                        title = "Important Variables: Contributions 3 * 2", 
                        col.points = col4ImportantVar23,
                        col.labels = col4ImportantVar23,
                        alpha.points = 0.5,
                        cex = 2.5, 
                        alpha.labels = 1, 
                        text.cex = 4,
                        font.face = "plain", 
                        font.family = "sans")
ctr.labels23 <- createxyLabels.gen(
  3,2, lambda = resMCA$ExPosition.Data$eigs,
  tau = resMCA$ExPosition.Data$t
)
a0009.Var.ctr23.imp  <- ctrV23.imp$zeMap  + ctr.labels23
#
print(a0009.Var.ctr23.imp)


## ----BR4var--------------------------------------------------------------
# Get the pseudo Bootstrap Ratios
BrLevels <- resMCA.inf$Inference.Data$fj.boots$tests$boot.ratios
wJ       <- 1 / resMCA.inf$Fixed.Data$ExPosition.Data$W
nIter    <- 1000
Br4Variables <- data4PCCAR::BR4varMCA(BrLevels, wJ, nIter) 
```

The variables identified in the heatmap as being important and confirmed with the 
contribution plots was all stable, indicated by the bootstrap ratios, expect for
Rock no being significant in the third dimension. It is called a Psedo Bootstrap 
Ratio because the variables are binned and are not exact. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----BR41----------------------------------------------------------------
VarBR1 <- Br4Variables$pseudoBR.pos[,1]
c0010.Var.br1  <- PrettyBarPlot2(VarBR1,
    main = 'Variable Pseudo Bootstrap Ratios: Dimension 1',
   ylim = c(0, 10),
    threshold = 2,
    font.size = 5,
    signifOnly = TRUE,
     color4bar = gplots::col2hex(color4Var)
)
print(c0010.Var.br1)

## ----BR42---------------------
VarBR2 <- Br4Variables$pseudoBR.pos[,2]
c0011.Var.br2  <- PrettyBarPlot2(VarBR2,
   main = 'Variable Pseudo Bootstrap Ratios: Dimension 2',
   ylim = c(0, 10),
  threshold = 2,
  font.size = 5,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(color4Var)
)
print(c0011.Var.br2)

## ----BR43------------------ 
VarBR3 <- Br4Variables$pseudoBR.pos[,3]
c0012.Var.br3  <- PrettyBarPlot2(VarBR3,
  main = 'Variable Pseudo Bootstrap Ratios: Dimension 3',
 ylim = c(0, 10),
  threshold = 3,
  font.size = 5,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(color4Var)
)
print(c0012.Var.br3)

```


Here the binned data is geometrically projected in the first two dimensions. 
It forms somewhat of an arch with the many of the low-end binned variables, indicating
low preference and low agreement, being in the third quadrant, and the high-end binned
variables, indicating high preference and high agreement, being in the first and fourth quadrant. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----createFjMap---------------------------------------------------------
axis1 = 1
axis2 = 2
Fj <- resMCA$ExPosition.Data$fj
# generate the set of maps
BaseMap.Fj <- createFactorMap(X = Fj , # J- Factor scores,
                              axis1 = axis1, axis2 = axis2,
                              title = 'MCA. Variables', 
                              col.points = col4Labels, 
                              cex = 1,
                              col.labels = col4Labels, 
                              text.cex = 2.5,
                              force = 2)
# add labels
labels4MCA <- createxyLabels.gen(x_axis = axis1, 
                                 y_axis = axis2,
               lambda = resMCA$ExPosition.Data$eigs,
               tau = resMCA$ExPosition.Data$t)
# make the maps
b0002.BaseMap.Fj <- BaseMap.Fj$zeMap + labels4MCA 
b0003.BaseMapNoDot.Fj  <- BaseMap.Fj$zeMap_background +
                         BaseMap.Fj$zeMap_text + labels4MCA 

## ----basemap Fj ----
print(b0002.BaseMap.Fj)
```


Classical, Swing, Alternative, and Rock and Roll are split by the first dimension, 
which makes sense as the first dimension accounts for 54% of the variance. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----mapJ-grey --------------------
col4Levels.imp <- data4PCCAR::coloringLevels(rownames(Fj),
                       col4ImportantVar)
BaseMap.Fj.imp <- createFactorMap(X = Fj, # Fj
                  axis1 = axis1, axis2 = axis2,
                  title = 'MCA. Important Variables', 
                  col.points = col4Levels.imp$color4Levels, 
                  cex = 1,
                  col.labels = col4Levels.imp$color4Levels, 
                  text.cex = 2.5,
                  force = 2)

b0010.BaseMap.Fj <- BaseMap.Fj.imp$zeMap + labels4MCA 
print(b0010.BaseMap.Fj)

```

Latino has an interesting pattern. If you have a high preference for Latino, you
tend to have a mid-range preference for Alternative, Metal, Opera, and Rock and Roll. 
If you don't like Latino, you may not like Classical or Swing either. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}

## ----adLines--------------------
lines4J <- addLines4MCA(Fj, 
                  col4Var = col4Levels.imp$color4Variables, 
                  size = .7)
 b0020.BaseMap.Fj <-  b0010.BaseMap.Fj + lines4J
 print( b0020.BaseMap.Fj)
 
```

 
```{r,echo=FALSE, warning=FALSE,message=FALSE}

## ----someLines -------------------------
zeNames          <- getVarNames(rownames(Fj)) 
importantsLabels <- zeNames$stripedNames %in% zeNames$variableNames[importantVar]
Fj.imp <- Fj[importantsLabels,]
lines4J.imp <- addLines4MCA(Fj.imp, 
            col4Var = col4Levels$color4Variables[which(importantVar)], 
            size = .9, linetype = 3, alpha = .5)
 b0021.BaseMap.Fj <-  b0020.BaseMap.Fj + lines4J.imp
 #print( b0021.BaseMap.Fj)
```


Latino has a similar pattern when projected in the second and third dimension, 
interacting with many other important contributors to these dimensions. It is mainly
separated by the third dimension, which compliments earlier analysis. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----mapJ23-grey-----------------------
col4Levels23.imp <- data4PCCAR::coloringLevels(rownames(Fj),
                                     col4ImportantVar23)
axis3 = 3
BaseMap.Fj23.imp <- createFactorMap(X = Fj , # resMCA$ExPosition.Data$fj,
                              axis1 = axis3, axis2 = axis2,
                              title = 'MCA. Important Variables. Dimensions 2 & 3',
                      col.points = col4Levels23.imp$color4Levels, 
                              cex = 1,
                      col.labels = col4Levels23.imp$color4Levels, 
                              text.cex = 2.5,
                              force = 2)
labels4MCA23 <- createxyLabels.gen(x_axis = 3, y_axis = 2,
               lambda = resMCA$ExPosition.Data$eigs,
               tau = resMCA$ExPosition.Data$t) 
b0030.BaseMap.Fj23 <- BaseMap.Fj23.imp$zeMap + labels4MCA23 

# zeNames          <- getVarNames(rownames(Fj)) 
importantsLabels23 <- zeNames$stripedNames %in% zeNames$variableNames[importantVar23]
Fj23.imp <- Fj[importantsLabels23,]
lines4J23.imp <- addLines4MCA(Fj23.imp, 
                    col4Var = col4Levels$color4Variables[
                               which(importantVar23)],
                    axis_h = axis3,
                    axis_v = axis2,
                    size = .9, linetype = 3, alpha = .5)
 b0031.BaseMap.Fj23 <-  b0030.BaseMap.Fj23 + lines4J23.imp
 print( b0031.BaseMap.Fj23)
```

```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----mapvarSup---------------
col4VarSup <- prettyGraphs::prettyGraphsColorSelection(
                                  ncol(cleanData.varSup))
Fj.sup <- resMCA.varSup$fjj
col4Levels.sup <- data4PCCAR::coloringLevels(
                      rownames(Fj.sup), col4VarSup)
BaseMap.Fj.sup <- createFactorMap(X = Fj.sup , # 
                axis1 = axis1, axis2 = axis2,
                constraints  = BaseMap.Fj$constraints, 
                # to get same size
      title = 'MCA. Supplementary and Important Variables', 
                col.points = col4Levels.sup$color4Levels, 
                              cex = 1,
                col.labels = col4Levels.sup$color4Levels, 
                text.cex = 2.5,
                force = 2)
lines4J.sup <- addLines4MCA(Fj.sup, 
                  col4Var = col4Levels.sup$color4Variables,
                  size = .7)
b0030.Sup.Fj <- BaseMap.Fj.sup$zeMap + 
                     BaseMap.Fj.imp$zeMap_dots + 
                     BaseMap.Fj.imp$zeMap_text +
                     labels4MCA + 
                     lines4J + lines4J.sup
#print(b0030.Sup.Fj)
```

The Bootstrap ratio can be sub-categorized for each binned variables. For example, 
Latino was binned from 1 to 5 with 5 being a high preference for Latino music. The 
bootstrap ratio for Latino 1, Latino 2, etc. can be computed. The analysis for the binned 
variables being geometrically projected across the first three dimension correspond 
with stable binned variables in the pseduo-bootstrap ratios. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----mapvarSup.only------------------------------------------------------
b0031.Sup.Fj.only <- BaseMap.Fj.sup$zeMap + 
                     BaseMap.Fj.imp$zeMap_dots + 
                     labels4MCA + 
                      lines4J.sup
#print(b0031.Sup.Fj.only)

## ----BR1-----------------------------------------------------------------

c0001.Levels.BR  <- PrettyBarPlot2(
  resMCA.inf$Inference.Data$fj.boots$tests$boot.ratios[,1], # BR
  main = 'Bootstrap Ratios for Columns : Dimension 1',
                             threshold = 2,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4Labels)
)
print(c0001.Levels.BR)

c0001.Levels.BR  <- PrettyBarPlot2(
  resMCA.inf$Inference.Data$fj.boots$tests$boot.ratios[,2], # BR
  main = 'Bootstrap Ratios for Columns : Dimension 2',
                             threshold = 2,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4Labels)
)
print(c0001.Levels.BR)

c0001.Levels.BR  <- PrettyBarPlot2(
  resMCA.inf$Inference.Data$fj.boots$tests$boot.ratios[,3], # BR
  main = 'Bootstrap Ratios for Columns : Dimension 3',
                             threshold = 2,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4Labels)
)
print(c0001.Levels.BR)
```

If we look at the distribution color with gender, it is heterogeneous with blue 
(males) forming an umbrella shape and being lower than the pink (females).


```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----createFiMap -----------------
Fi <- resMCA$ExPosition.Data$fi
colCity <- c('deeppink', 'blue')
nI <- nrow(Fi)
col4I.City <- rep("",nI)

for (i in 1:length(colCity) ){
  lindex <- cleanData.allVar[,'Gender'] %in% unique(cleanData.allVar[,'Gender'])[i]
  col4I.City[lindex] <- colCity[i]
}


# generate the set of maps
BaseMap.Fi <- createFactorMap(X = Fi , # resMCA$ExPosition.Data$fj,
                        axis1 = axis1, axis2 = axis2,
                        title = 'MCA. Observations (by Gender)', 
                        col.points = col4I.City,
                        alpha.points = .4, cex = .9,
                        col.labels = col4I.City,
                        text.cex = 2.5, 
                        force = 2)
# make the maps
d0001.BaseMapNoLabels.Fi  <- BaseMap.Fi$zeMap_background +
                                 BaseMap.Fi$zeMap_dots + 
                                 labels4MCA 

## ----plotaMapi, fig.width= 8---------------------------------------------
print(d0001.BaseMapNoLabels.Fi)

```

The Bootstrap interval allows us to form a 95% of resampling map around the mean 
the specific gender. We see that slightly overlap and do not form distinct group. 
So while the plot of the observation look heterogeneous, it is statistically 
homogeneous. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}

## ----Boot4CI-------------------------------------------------------------
# Bootstrap for CI:
BootCube.Gr <- PTCA4CATA::Boot4Mean(resMCA$ExPosition.Data$fi, 
                                 design = cleanData.allVar$Gender,
                                 niter = 100,
                                 suppressProgressBar = TRUE)
nationsMeans <- PTCA4CATA::getMeans(resMCA$ExPosition.Data$fi, cleanData.allVar$Gender)
# colCity <- c('darkblue', 'red4')
MapGroup <- PTCA4CATA::createFactorMap(nationsMeans,
                            # use the constraint from the main map
                            constraints = BaseMap.Fi$constraints,
                            col.points = colCity,
                            cex = 7,  # size of the dot (bigger)
                            col.labels = colCity,
                            text.cex = 6)
d002.Map.I.withMeans <- d0001.BaseMapNoLabels.Fi  +
                          MapGroup$zeMap_dots + MapGroup$zeMap_text
print(d002.Map.I.withMeans)

```

```{r, echo=FALSE, warning=FALSE,message=FALSE}

## ----graphElli-----------------------------------------------------------
GraphElli <- PTCA4CATA::MakeCIEllipses(BootCube.Gr$BootCube[,1:2,],
                            names.of.factors = c("Dimension 1","Dimension 2"),
                            col = colCity,
                            p.level = .95)
d003.Map.I.withCI <-  d0001.BaseMapNoLabels.Fi + 
                          MapGroup$zeMap_text +  GraphElli
print(d003.Map.I.withCI)

## ----TI------------------------------------------------------------------
GraphTI.Hull <- PTCA4CATA::MakeToleranceIntervals(resMCA$ExPosition.Data$fi,
                            design = as.factor(cleanData.allVar$Gender),
                            # line below is needed
                            names.of.factors =  c("Dim1","Dim2"), # needed 
                            col = colCity,
                            line.size = .50, 
                            line.type = 3,
                            alpha.ellipse = .2,
                            alpha.line    = .4,
                            p.level       = .75)
#_____________________________________________________________________
# Create the map:
d005.Map.I.withTIHull <- d002.Map.I.withMeans  +
                           GraphTI.Hull + MapGroup$zeMap_dots +
                           MapGroup$zeMap_text + MapGroup$zeMap_dots
#_____________________________________________________________________
```


As stated earlier, the males tend to congregate in the third quadrant more than females. 
Some males also tend to report lower rating on Classical, Swing, and Latino. Thus, 
the tolerance interval gives a better visualization, via  a convex hull, for how 
some males report their music preferences in comparison to females. However, 
this should be view with caution as the confidence intervals overlap between 
types of gender. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
# plot it

print(d005.Map.I.withTIHull)
```

## Summary

The Scree plot shows that there are three dimensions of interest, at least. 
Classical,Swing, and Rock n Roll contribute the most to dimension 1. Dance, 
Pop, Hiphop, and Latino contribute the most to dimension 2. Fast, Punk, 
Classical, and Opera contribute the most to dimension 3. 

In conclusion, Dance, Pop, Techo, Latino, and Hiphop form there own 
group in relation to the other music groups. Bootstraps for dimension 1 were 
significant. Based on the Important Variables MCA plot, Latino is related to 
Punk and Rock n Roll. Gender means overlap, indication that the distinction 
between groups is weak. However, the MCA Observations (by Gender) plot 
shows a Convex Hull. 
