
# Barycentric Discriminant Analysis {#bada}

This method is a robust version of discriminant analysis, which groups observations into 
pre-defined groups such as COVID-19 positive or negative, employed or unemployed, or 
married, divorced, separated, or single. BADA can even be used when n << p. 


```{r, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE}
rm(list = ls())
graphics.off()
gc()
```

```{r, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE}
#leTitre   <- 'BADA: Music and Education'
#leDir     <- paste0(getwd(),'/')    # Where am I
#filename  <- 'Music&Education' # my title
#path2save <-  paste0(leDir, filename)
```

```{r, include = FALSE}
library(tidyverse)
library(ExPosition)
#devtools::install_github("LukeMoraglia/ExPosition1/TExPosition")
#devtools::install_github("LukeMoraglia/ExPosition1/TInPosition")
library(TExPosition)
library(TInPosition)
library(PTCA4CATA)
# devtools::install_github('HerveAbdi/data4PCCAR')
library(data4PCCAR)
library(corrplot)
library(kableExtra)
library(dplyr)
```

## Cleaning the data

```{r, echo=FALSE, warning=FALSE,message=FALSE}
# Load data
music = read.csv("data/music10.csv")

data_2=music[,-1]
rownames(data_2)=music[,1]
data_3 = na.omit(data_2)
rawData <- data_3
descriptors <-  rawData[,1:2]
Xmat <-  rawData[, 3:19]
Ymat <-  rawData[, 20]
XYmat <- rawData[, 1:19]
colnames(XYmat) <- c('Music','Fast','Dance', 'Folk','Country','Classical','Musical',
                        'Pop','Rock','Metal','Punk','Hiphop','Reg','Swing','RnR','Alt',
                        'Lat','TT','Opera')
```


The data set is the same from PCA, evaluating music preferences in 931 
participants. 
```{r, warning=FALSE,message=FALSE}

head(rawData) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```

## Heat map


The heat map is on a matrix with the groups of education level on the 
rows, the variables of music preference on the columns, and the average survey
score for the respective variables for each of the education levels for the table 
values. 
```{r, echo=FALSE, warning=FALSE,message=FALSE}
#heatmap
XYmat2 = as.matrix(XYmat)
Ymat2 = as.matrix(Ymat)

CorrMeans = PTCA4CATA::getMeans(XYmat2, Ymat2)

CorrMeans2 = as.matrix(CorrMeans)
corrplot3 = corrplot((CorrMeans2), 
                          method = "shade",
                          col = colorRampPalette(c("white", "blue"))(200),
                          tl.pos = "lt",
                           tl.col = "black",
                           tl.cex = 0.8,
                           is.corr = FALSE,
                           addCoefasPercent = TRUE,
                           number.cex = 0.5)

```

## Analysis

```{r, warning=FALSE,message=FALSE, results=FALSE}
#_________________________________________________
# Computations ----
# Run BADA  ----
resBADA <- tepBADA(XYmat, DESIGN = Ymat,
                   graphs = FALSE)
# Inferences ----
#set.seed(70301) # we had a problem 
# with the inference part
# it is addressed iin the Fix from Luke's github
nIter = 889
resBADA.inf <- tepBADA.inference.battery(XYmat, 
                  DESIGN = Ymat,
                  test.iters = nIter,
                  graphs = FALSE)

#_________________________________________________
#_________________________________________________
```


In the plot of the observations and the barycenter means, we see doctoral degree
and currently a primary school student are distict groups from each other and the 
other education levels. The rest of the means are near the center of gravity. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
# Graphics ----
# scree for ev  ----
# The ScreePlot. Fixed Effects. ----
# Get the ScreePlot
# scree for ev ----
PlotScree(ev = resBADA$TExPosition.Data$eigs,
   title = 'BADA Education Level and Music: Inertia Scree Plot',
   plotKaiser = FALSE, 
   color4Kaiser = ggplot2::alpha('darkorchid4', .5),
   lwd4Kaiser  = 2)
# Save the plot for pptx
a0002.Scree.sv <- recordPlot()
#_________________________________________________
# wineColors ----
wineColors <- Ymat
# get colors for graphs
wineColors <- recode(wineColors, 
                     "primary school" = 'blue', 
            "currently a primary school pupil" = 'gold', 
            "secondary school" = 'lightpink2',
            "college/bachelor degree" = 'green',
            "masters degree" = 'orange',
            "doctorate degree" = 'red')
#_________________________________________________
#  Observations and means ----
# Observations ----
#_________________________________________________
# I-set map ----
# a graph of the observations
Imap <- PTCA4CATA::createFactorMap(
  resBADA$TExPosition.Data$fii,
  cex = 1, 
   display.labels = FALSE,
  col.points = wineColors,
  col.labels = wineColors,
  alpha.points = .5
)

#_________________________________________________
# make labels ----
label4Map <- createxyLabels.gen(1,2,
          lambda = resBADA$TExPosition.Data$eigs,
          tau = resBADA$TExPosition.Data$t)
#_________________________________________________
# Fix Fi ----
# Note that current version of tepBADA 
# has a scaling problem
# Will soon be fixed
# wineMeans <- resBADA$TExPosition.Data$fi 
# code when BQDQ is fixed
wineMeans <- PTCA4CATA::getMeans(
               resBADA$TExPosition.Data$fii, 
               Ymat)

# a vector of color for the means
# Explicit recoding to make sure 
# that the names match
# => Needs to be improved!
col4Means <- recode(rownames(wineMeans), 
                    "primary school" = 'blue', 
                    "currently a primary school pupil" = "pink",
                     "secondary school" = 'brown', 
                     "college/bachelor degree" = 'orange',
                     "masters degree" = "green",
                     "doctorate degree" = "red")
names(col4Means) <- rownames(wineMeans)
# the map
MapGroup <- PTCA4CATA::createFactorMap(wineMeans,
             # use the constraint from the main map
             constraints = Imap$constraints,
             col.points = col4Means,
             cex = 4,  # size of the dot (bigger)
              display.labels = TRUE,
             col.labels = col4Means,
             text.cex = 3)

# The map with observations and group means
a003.bada <- Imap$zeMap + 
               label4Map +
               MapGroup$zeMap_dots + 
               MapGroup$zeMap_text
print(a003.bada)
```


The confidence intervals confirm that currently a primary student is different from 
all the other education levels as is doctoral degree from all the other education 
levels. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#_________________________________________________
# Confidence intervals
# Bootstrapped CI ----
#_________________________________________________
# Create Confidence Interval Plots
fi.boot <- resBADA.inf$Inference.Data$boot.data$fi.boot.data$boots
# We want to use the rownames of fi.boot 
# as reference to get the correct
# color. 
# However, the original rownames include "." 
# and don't match with 
# the original row names. 
# So, the `sub` function was used to get rid of 
# the "." by replacing all "." 
# in the rownames of fi.boot as an empty 
# string.
rownames(fi.boot) <- sub("[[:punct:]]","",
                          rownames(fi.boot))
# use function MakeCIEllipses 
# from package PTCA4CATA
GraphElli <- PTCA4CATA::MakeCIEllipses(
    resBADA.inf$Inference.Data$boot.data$fi.boot.data$boots,
    col = col4Means[rownames(fi.boot)], 
    # use rownames as reference to pick the color
    p.level = .95
)
#_________________________________________________
# create the I-map with Observations, 
# means and confidence intervals
#
a004.bada.withCI <-  Imap$zeMap_background  + 
                        Imap$zeMap_dots     + 
                        MapGroup$zeMap_dots + 
                        MapGroup$zeMap_text +
                      GraphElli + label4Map +
  ggtitle('BADA: Group Centers with CI and Observations')
#_________________________________________________
# plot it!
#dev.new()
print(a004.bada.withCI)
```


Here we see that Masters degree is different from doctoral degree, secondary school,
and currently a primary school student. Secondary school, likewise, is different 
from doctoral degree, masters degree, and currently a primary school student. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
# create the I-map with Observations, 
# means and confidence intervals
#
a004.bada.withCI <-  Imap$zeMap_background  + 
                       
                        MapGroup$zeMap_text +
                      GraphElli + label4Map +
  ggtitle('BADA: Group Centers with CI and Observations')
#_________________________________________________
# plot it!
#dev.new()
print(a004.bada.withCI)
```


The tolerance intervals shows the the spread of 95% of doctoral degree observations 
overlap with the groups near the center of gravity as does the spread of 95% of 
currently a primary school student. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#_________________________________________________
# with Hull ----
Ymat3 = as.factor(Ymat)
Fii <- resBADA$TExPosition.Data$fii

colnames(Fii) <- paste0('Dimension ', 1:ncol(Fii))

# getting the color correct: an ugly trick
col4Hull <- col4Means[match(names(col4Means), 
                     levels(Ymat3) )]

#is.data.frame(Ymat)

#dim(Ymat3)
#dim(Fii)

GraphHull <- PTCA4CATA::MakeToleranceIntervals(
                      Fii,
                      design = Ymat3,
                      col = col4Hull,
                # the next line is required 
                # for some strange unknown reasons
          names.of.factors =  c("Dim1","Dim2"),
          p.level = .66)



#
a006.bada.withHull <-  (Imap$zeMap_background  + 
                          Imap$zeMap_dots     + 
                          MapGroup$zeMap_dots + 
                          MapGroup$zeMap_text +
                          GraphHull           + 
                          label4Map           +
    ggtitle('BADA: Group Centers with Hulls and Observations'))
a006.bada.withHull
```


The loading show that doctoral students tend to like Opera, Folk, Latino, Metal, 
and Alternative while currently a primary student individual prefer Swing, Punk, 
Hiphop, and Country. 



```{r, echo=FALSE, warning=FALSE,message=FALSE}
#_________________________________________________
# J-set ----
# gt colors
col4X <- prettyGraphsColorSelection(
                   n.colors = ncol(descriptors),
                   starting.color = 42)

col4Y <- prettyGraphsColorSelection(
                    n.colors = ncol(Xmat),
                    starting.color = 13)
col4Var = c(col4X,col4Y)

#_________________________________________________
Fj <- resBADA$TExPosition.Data$fj
baseMap.j <- PTCA4CATA::createFactorMap(
                        Fj,
                        col.points   = col4Var,
                        alpha.points =  .3,
                        col.labels   = col4Var)

#_________________________________________________
# arrows
zeArrows <- addArrows(Fj, color = col4Var)

# A graph for the J-set
# A graph for the J-set
b001.aggMap.j <- baseMap.j$zeMap_background + # background layer
                      baseMap.j$zeMap_dots + 
                      baseMap.j$zeMap_text +  # dots & labels
                      label4Map 
b002.aggMap.j <- b001.aggMap.j + zeArrows
# We print this Map with the following code
#dev.new()
print(b002.aggMap.j)
```


The contribution show that Latino, Alternative, Swing, Punk, Hiphop, and Country 
are important contributors to the first two dimensions while the bootstrap ratios
show that Latino, Alternative, Swing, Hiphop, and Country and stable variables. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
#_________________________________________________
#  Contributions ####
#_________________________________________________
#_________________________________________________
# Ctr J-set 
###### 1 ====
# 
ctrj <- resBADA$TExPosition.Data$cj
signed.ctrj <- ctrj * sign(Fj)
# BR1
c001.plotCtrj.1 <- PrettyBarPlot2(
           bootratio = round(100*signed.ctrj[,1]), 
           threshold = 100 / nrow(signed.ctrj), 
           ylim = NULL,
           signifOnly = TRUE,
           color4bar = gplots::col2hex(col4Var),
           color4ns = "gray75", 
           plotnames = TRUE, 
    main = 'Important Contributions Variables. Dim 1.', 
           ylab = "Signed Contributions")
#dev.new()
print(c001.plotCtrj.1)
#_________________________________________________
###### 2 ====
# 
c002.plotCtrj.2 <- PrettyBarPlot2(
  bootratio = round(100*signed.ctrj[,2]), 
  threshold = 100 / nrow(signed.ctrj), 
  ylim = NULL, 
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4Var),
  color4ns = "gray75", 
  plotnames = TRUE, 
  main = 'Important Contributions Variables. Dim 2.', 
  ylab = "Signed Contributions")
#dev.new()
print(c002.plotCtrj.2)
#_________________________________________________
#_________________________________________________
#  Bootstrap ratios ####
#_________________________________________________
#_________________________________________________
#BR. 1 ====
# 
BRj <- resBADA.inf$Inference.Data$boot.data$fj.boot.data$tests$boot.ratios
# BR1
d001.plotBRj.1 <- PrettyBarPlot2(
  bootratio = BRj[,1], 
  threshold = 2, 
  ylim = c(-1,4), 
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4Var),
  color4ns = "gray75", 
  plotnames = TRUE, 
  main = 'Bootstrap Ratios Variables. Dim 1.', 
  ylab = "Bootstrap Ratios")
#dev.new()
print(d001.plotBRj.1)
#_________________________________________________
###### 2 ====
# 
d003.plotBRj.2 <- PrettyBarPlot2(
  bootratio = BRj[,2], 
  threshold = 2, 
  ylim = c(-3,1), 
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4Var),
  color4ns = "gray75", 
  plotnames = TRUE, 
  main = 'Bootstrap Ratios Variables. Dim 2.', 
  ylab = "Bootstrap Ratios")
#dev.new()
print(d003.plotBRj.2)

```

## Fixed Confusion Matrix


According to chance, their is a 0.1667 or 16.67% likelihood to discriminant to 
the barycenter groups. The fixed confusion matrix predicted at a accuracy of 0.20,
which is more than by change. Thus, the level of education is a modest benchmark to
classify music preference. The horizontal titles are the actual numbers while the 
vertical titles are the predicted numbers. The diagonal from the upper left to 
the bottom right represents the hits while the other values represent the misses. 

```{r}
fixed_cm <-resBADA.inf$Inference.Data$loo.data$fixed.confuse

head(fixed_cm) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

resBADA.inf$Inference.Data$loo.data$fixed.acc

                            
                          
```

## Random Confusion Matrix


The Random Confusion Matrix only predicted by 0.1664, which under-performed compared
to the liklihood by chance alone. The Random Confusion Matrix is a method to 
mimic training and testing dat sets. It utilizes Jackknife whereby one observation is
left out, and a permutation test is run. The BADA is performed on the Leave One Out
Data Set and the left out data point is used to minimize the distance to barycenter. 
Thus, random confusion matrices are a way of validating the fixed confusion matrix
through Jackknife and permutation testing. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
random_cm <- resBADA.inf$Inference.Data$loo.data$loo.confuse

head(random_cm) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

resBADA.inf$Inference.Data$loo.data$loo.acc

```

## Summary

The Scree Plot shows that the first three dimension are important. The main take away from the confidence intervals plot is that currently a primary school student and doctoral student are different from each other and all other groups. However, the random confusion matrix showed a poor accuracy in the prediction strength of the BADA algorithm. 
