
# Discriminant Correspondence Analysis {#dica}

This method is an extension of Discriminant Analysis and Correspondence Analysis,
with the caveat of containing nominal variables for the pre-defined groups. Traditionally, a
comparison between a training data set and testing data set is done to evaluate the 
effectiveness of the classification ability of the analysis.  

```{r, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE}
# Clean start ----
rm(list = ls())
graphics.off()
gc()
```

```{r, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE}
#_________________________________________________
# parameters for pptx ----
#leTitre   <- 'DICA: Music and Education' # my title
#leDir     <-  paste0(getwd(),'/')   # Where I am
#filename  <- 'DICA4Music&Education' 
               # file name for results
#path2save <-  paste0(leDir, filename)
#_________________________________________________
#_________________________________________________
# Preamble ----
## libraries ----
#install.packages("Ckmeans.1d.dp")
# Make sure to install packages last versions. 
library(tidyverse)
library(ExPosition)
#install.packages('TInPosition') # if needed
library(TExPosition)
library(TInPosition)
#devtools::install_github('HerveAbdi/PTCA4CATA')
library(PTCA4CATA)
#devtools::install_github('HerveAbdi/data4PCCAR')
library(data4PCCAR)
library(Ckmeans.1d.dp) # to cut 
library(corrplot)
library(printr)    # To pretty print tables 
library(kableExtra)
```

This is a for loop that is used for the color designation later in this method. 


```{r}
## foo function ----
# to be used later 
foo <- function(x) {
  xuniq <- unique(x)
  N <- length(xuniq)
  res <- rep(NA, N)
  for (i in 1:N) {
    res[i] <- sum(x == xuniq[i])
  }
  return(res)
}
```

```{r, echo=FALSE, warning=FALSE,message=FALSE}
# The data ----
music = read.csv("data/music11.csv")
data_2=music[,-1]
rownames(data_2)=music[,1]

data_3 = na.omit(data_2)
rawData <- data_3


```

I binned the variables just like in MCA, but I did not bin the discriminant varibale of 
alcohol consumption. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
## recode as factors ----
# Here we bin the data.
## bin the data ---- 

colnames(rawData) <- c('Music','Fast','Dance', 'Folk','Country','Classical','Musical',
                        'Pop','Rock','Metal','Punk','Hiphop','Reg','Swing','RnR','Alt',
                        'Lat','TT','Opera','Ed')


rawData[,'Music'] <- plyr::mapvalues(rawData[,'Music'], 
                                   from = c("1", "2","3","4", "5"), to = c("1","1","1","1", "2"))
rawData[,'Fast'] <- plyr::mapvalues(rawData[,'Fast'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","1","2","3", "3"))
rawData[,'Dance'] <- plyr::mapvalues(rawData[,'Dance'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","1","2","3", "3"))
rawData[,'Folk'] <- plyr::mapvalues(rawData[,'Folk'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","3", "3"))
rawData[,'Country'] <- plyr::mapvalues(rawData[,'Country'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","3", "3"))
rawData[,'Classical'] <- plyr::mapvalues(rawData[,'Classical'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "5"))
rawData[,'Musical'] <- plyr::mapvalues(rawData[,'Musical'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "4"))
rawData[,'Pop'] <- plyr::mapvalues(rawData[,'Pop'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","1","2","3", "4"))
rawData[,'Rock'] <- plyr::mapvalues(rawData[,'Rock'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","1","1","2", "3"))
rawData[,'Metal'] <- plyr::mapvalues(rawData[,'Metal'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","3", "3"))
rawData[,'Punk'] <- plyr::mapvalues(rawData[,'Punk'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "4"))
rawData[,'Hiphop'] <- plyr::mapvalues(rawData[,'Hiphop'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "5"))
rawData[,'Reg'] <- plyr::mapvalues(rawData[,'Reg'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "4"))
rawData[,'Swing'] <- plyr::mapvalues(rawData[,'Swing'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "4"))
rawData[,'RnR'] <- plyr::mapvalues(rawData[,'RnR'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","1","2","3", "4"))
rawData[,'Alt'] <- plyr::mapvalues(rawData[,'Alt'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "5"))
rawData[,'Lat'] <- plyr::mapvalues(rawData[,'Lat'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","4", "5"))
rawData[,'TT'] <- plyr::mapvalues(rawData[,'TT'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","3", "3"))
rawData[,'Opera'] <- plyr::mapvalues(rawData[,'Opera'], 
                                      from = c("1", "2","3","4", "5"), to = c("1","2","3","3", "3"))

#rawData[,'Ed'] <- plyr::mapvalues(rawData[,'Ed'], 
                                      #from = c("primary school","currently a primary school pupil","secondary school","college/bachelor degree","masters degree","doctorate degree"), to = c("P","C","S","B", "M", "D"))

#rawData[,20] <- as.numeric(rawData[,20])
#hist.E <- hist(rawData[,20], breaks = 20) # 1,2,4,5,6 and 3

descriptors <-  rawData[,1:2]
Xmat <-  rawData[, 3:19]
Ymat <-  rawData[, 20]

XYmat <- rawData[, 1:19]
XYmat[,1:19] <- lapply(XYmat[,1:19], as.integer)
```

```{r}

head(XYmat) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```

### Heatmap


```{r, echo=FALSE, warning=FALSE,message=FALSE}
#heatmap
XYmat_disjunct = makeNominalData(XYmat)

group_sums = getMeans(XYmat_disjunct, Ymat, FUN = sum)


corrplot(as.matrix(group_sums), is.corr = FALSE, method = "color", col = 
           colorRampPalette(c("white", "darkblue"))(200), 
         tl.cex = .5, cl.cex = .3)
row_profile = group_sums/rowSums(group_sums)
```

```{r, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE }
rowSums(row_profile)
```

```{r, echo=FALSE, warning=FALSE,message=FALSE}
corrplot(as.matrix(row_profile), is.corr = FALSE, method = "color", col = 
       colorRampPalette(c("white", "darkblue"))(200), tl.cex = .5, cl.cex = .3)
```


## Analysis



```{r, warning=FALSE,message=FALSE, results=FALSE}
#_________________________________________________
# Computations ----
## Run DiCA  ----
resDiCA <- tepDICA(XYmat, 
                   make_data_nominal = TRUE, 
                   DESIGN = Ymat,
                   graphs = FALSE)


##  Inferences ----
set.seed(70301) # set the seed
# for random so that we all have the same results. 
nIter <- 100

resDiCA.inf <- tepDICA.inference.battery(XYmat,
                 DESIGN = Ymat,
                 test.iters = nIter,
                 graphs = FALSE)

```

The Scree plot reveals that the there are two dimensions with the first capturing 
almost 70% of the total variance and the second capturing the remaining amount. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#_________________________________________________
#_________________________________________________
# Graphics ----
##  scree. Fixed  ----
## Fixed Effects. 
PlotScree(ev = resDiCA$TExPosition.Data$eigs,
  title = 'DICA Music and Education',
  plotKaiser = FALSE, 
  color4Kaiser = ggplot2::alpha('darkorchid4', .5),
  lwd4Kaiser  = 2)
# Save the plot
a0002.Scree.sv <- recordPlot()
#_________________________________________________
##  EdColors ----
AlColors <- Ymat

AlColors <- dplyr::recode(AlColors, 
            "never" = 'blue', 
            "social drinker" = 'deeppink',
            "drink a lot" = "green")
       
         
#_________________________________________________
##  Observations and means ----
###   Observations ----
#_________________________________________________
### I-set map ----
# a graph of the observations
Imap <- PTCA4CATA::createFactorMap(
  resDiCA$TExPosition.Data$fii,
  display.labels = FALSE,
  col.points = AlColors,
  col.labels = AlColors,
  cex = 1,
  alpha.points = .2
)
```


The group means for the amount of alcohol consumed are distinct: Never drinks
is separate from the social drinkers, which both are separate from the drinks a lot
category. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
#_________________________________________________
### make labels ----
label4Map <- createxyLabels.gen(1,2,
           lambda = resDiCA$TExPosition.Data$eigs,
           tau = resDiCA$TExPosition.Data$t)
#______________________________________________________________
### ** Fix Fi means ** ----
# wineMeans <- resDiCA$TExPosition.Data$fi 
# code when BADA is fixed
EdMeans <- PTCA4CATA::getMeans(
                    resDiCA$TExPosition.Data$fii, 
                    Ymat)
# a vector of color for the means
# Explicit recoding to make sure the names match
# Needs to be improved
col4Means <- recode(rownames(EdMeans), 
                  "never" = 'blue', 
            "social drinker" = 'deeppink',
            "drink a lot" = "green")
### the map ----
MapGroup <- PTCA4CATA::createFactorMap(EdMeans,
              display.labels = TRUE,
              constraints = Imap$constraints, # use the constraints from the main map
              col.points = col4Means,
              cex = 4,  # size of the dot (bigger)
              col.labels = col4Means,
              text.cex = 4)
# The map with observations and group means
a003.DICA <- Imap$zeMap + label4Map +
    MapGroup$zeMap_dots + MapGroup$zeMap_text
print(a003.DICA)
```

The confidence intervals confirm that the group means are distinct from each other. 



```{r, warning = FALSE, echo=FALSE, message=FALSE, results = FALSE}
#_________________________________________________
# Confidence intervals
### Bootstrapped CI ----
#_________________________________________________
# Create Confidence Interval Plots
# use function MakeCIEllipses 
# from package PTCA4CATA
# First get the order of colors for the ellipses
truc <- unique(rownames(
   resDiCA.inf$Inference.Data$boot.data$fi.boot.data$boots))
col4Means.ordered <- col4Means[order(truc)]
#
GraphElli <- PTCA4CATA::MakeCIEllipses(
  resDiCA.inf$Inference.Data$boot.data$fi.boot.data$boots,
  col = col4Means.ordered, 
  centers = resDiCA$TExPosition.Data$fi,
  p.level = .95
)

#_________________________________________________
### Pretty Imap ----
# create the I-map with Observations, 
# means and confidence intervals
#
a004.DICA.withCI <- Imap$zeMap_background + 
                         Imap$zeMap_dots + 
                         MapGroup$zeMap_dots + 
                         GraphElli + label4Map +
  
  ggtitle('DICA: Group Centers with CI and Observations')
#_________________________________________________
# plot it!
# dev.new()
print(a004.DICA.withCI)
```

```{r,  warning = FALSE, echo=FALSE, message=FALSE, results = FALSE}
### Pretty Imap ----
# create the I-map with Observations, 
# means and confidence intervals
#
aa004.DICA.withCI <- Imap$zeMap_background + 
                       
                         GraphElli + label4Map +
  MapGroup$zeMap_text
  ggtitle('DICA: Group Centers with CI and Observations')
#_________________________________________________
# plot it!
# dev.new()
print(aa004.DICA.withCI)

```

The tolerance inteval, however, show that the spread of 95% of the observations 
overlap between the three categorical variables for alcohol consumption. 

```{r,  warning = FALSE, echo=FALSE, message=FALSE}
#_________________________________________________
###  with Hull ----
Fii <- resDiCA$TExPosition.Data$fii

# use function MakeToleranceIntervals 
#     from package PTCA4CATA
colnames(Fii) <- paste0('D', 1:2)
GraphHull <- PTCA4CATA::MakeToleranceIntervals(
                   Fii,
                   design = Ymat,
                   col = col4Means,
                   type = 'hull',
                   p.level = 1.0)
#(
a006.DICA.withHull <-  Imap$zeMap_background + 
                       Imap$zeMap_dots       + 
                       MapGroup$zeMap_dots   + 
                       MapGroup$zeMap_text   +
                       GraphHull + label4Map +
 ggtitle('DICA: Group Centers with Hulls and Observations')
#)
# To print the Hulls
#dev.new()
print(a006.DICA.withHull )

```


If you drink a lot you tend to like Metal and Swing more than others. If you are
a social drinker you tend to like Latino and Pop whereas if you have never consumed 
alcohol, you prefer Country and Swing. 


```{r,  warning = FALSE, echo=FALSE, message=FALSE}
#_________________________________________________
## Variables ----
### colors 4 J-set ----
# get colors
nominalcolnames <- gsub("\\.[0-9]", "", 
     drop(colnames(resDiCA$TExPosition.Data$X)))
col4Var <- prettyGraphsColorSelection(
                        n.colors = ncol(XYmat),
                        starting.color = 42)
col4VarNom <- rep(col4Var,foo(nominalcolnames))
#_________________________________________________
### base map J ----
Fj <- resDiCA$TExPosition.Data$fj
baseMap.j <- PTCA4CATA::createFactorMap(Fj,
                 col.points   = col4VarNom,
                 alpha.points =  .3,
                 col.labels   = col4VarNom)
#_________________________________________________
b001.BaseMap.Fj <- baseMap.j$zeMap + label4Map 
b002.BaseMapNoDot.Fj  <-
  baseMap.j$zeMap_background +
  baseMap.j$zeMap_text + label4Map 
###  add Lines ----
lines4J <- addLines4MCA(Fj, col4Var = col4Var)
b003.MapJ <-  b001.BaseMap.Fj + lines4J
#dev.new()
print(b003.MapJ)

```


As referenced in the above analysis, high ratings of Metal, Swing, and Pop are
important contributors to first dimension. Country is a important contributor to the 
second dimension. Latino and Pop are both significant bootstrap ratios. 

```{r,  warning = FALSE, echo=FALSE, message=FALSE}

#_________________________________________________
##  Contributions ----
#_________________________________________________
#_________________________________________________
### Ctr J-set ----
# 
ctrj <- resDiCA$TExPosition.Data$cj
signed.ctrj <- ctrj * sign(Fj)
####  CtrJ 1 ====
c001.plotCtrj.1 <- PrettyBarPlot2(
  bootratio = round(100*signed.ctrj[,1]), 
  threshold = 100 / nrow(signed.ctrj), 
  ylim = NULL, 
    signifOnly = TRUE,
  color4bar = gplots::col2hex(col4VarNom),
  color4ns = "gray75", 
  plotnames = TRUE, 
  main = 'Important Contributions Variables. Dim 1.', 
  ylab = "Signed Contributions")
#dev.new()
print(c001.plotCtrj.1)
#_________________________________________________
#### CtrJ 2 ====
c002.plotCtrj.2 <- PrettyBarPlot2(
  bootratio = round(100*signed.ctrj[,2]), 
  threshold = 100 / nrow(signed.ctrj), 
  ylim = NULL,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4VarNom),
  color4ns = "gray75", 
  plotnames = TRUE, 
  main = 'Important Contributions Variables. Dim 2.', 
  ylab = "Signed Contributions")
#dev.new()
print(c002.plotCtrj.2)
#_________________________________________________
### Contribution Maps ----
CtrJ12 <- data4PCCAR::ctr4Variables(
                     resDiCA$TExPosition.Data$cj)
baseMap.ctrj <- PTCA4CATA::createFactorMap(CtrJ12,
                     col.points   = col4Var,
                     alpha.points =  .3,
                     col.labels   = col4Var)
#_________________________________________________
b001a.BaseMap.Ctrj <- baseMap.ctrj$zeMap + 
                       label4Map +
  ggtitle('Variables Contributions Map')
b001aa.BaseMapNoDot.Ctrj  <- 
  baseMap.ctrj$zeMap_background +
  baseMap.ctrj$zeMap_text + label4Map 
#dev.new()
print(b001a.BaseMap.Ctrj)
#_________________________________________________
#_________________________________________________
##  Bootstrap ratios ----
#_________________________________________________
### BR4Variables ----
# Note that we do not have the BR for the levels, 
# only of the variables
#_________________________________________________
### BR. 1 ====
# 
BRj <- resDiCA.inf$Inference.Data$boot.data$fj.boot.data$tests$boot.ratios
# BR1
d001.plotBRj.1 <- PrettyBarPlot2(
  bootratio = BRj[,1], 
  threshold = 2, 
  ylim = NULL, 
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4Var),
  color4ns = "gray75", 
  plotnames = TRUE, 
  main = 'Bootstrap Ratios Variable Levels. Dim 1.', 
  ylab = "Bootstrap Ratios")
#dev.new()
print(d001.plotBRj.1)
#_________________________________________________
### BR. 2 ====
# 
d003.plotBRj.2 <- PrettyBarPlot2(
  bootratio = BRj[,2], 
  threshold = 2, 
  ylim = c(-1,3), 
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4Var),
  color4ns = "gray75", 
  plotnames = TRUE, 
  main = 'Bootstrap Ratios Variable Levels. Dim 2.', 
  ylab = "Bootstrap Ratios")
#dev.new()
print(d003.plotBRj.2)
```

## Fixed Confusion Matrix


The Fixed Confusion matrix performs slightly better than change (0.39 vs. 0.33).


```{r, echo=FALSE, warning=FALSE,message=FALSE}
fixed_cm <-resDiCA.inf$Inference.Data$loo.data$fixed.confuse
  
head(fixed_cm) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

resDiCA.inf$Inference.Data$loo.data$fixed.acc
```

## Random Confusion Matrix


The random confusion matrix also performs slightly better than chance (0.37 vs.
0.33).


```{r, echo=FALSE, warning=FALSE,message=FALSE}
random_cm <- resDiCA.inf$Inference.Data$loo.data$loo.confuse

head(random_cm) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

resDiCA.inf$Inference.Data$loo.data$loo.acc
```



## Summary

Discriminant Correspondence Analysis have feature of BADA and MCA with a emphasis
on predicting the closest distance to a group mean binned data. The analysis revealed
that heavy drinkers prefer Metal and Swing while social drinkers  tend to like 
Latino and Pop. Those that have never consumed alcohol prefer Country and Swing.
The fixed and random matrices preformed slightly better than chance. 
