
# Partial Least Square Correlation {#plsc}

Using information from the same observations, PLSC finds the correlation of 
multivariate data in two data tables. The first step is to obtain latent 
variables from linear combination similar to PCA. Analogously, these latent 
variables maximize the covariance between the tables. Additionally, factor 
scores in PCA are akin to latent variables in PLSC while loadings in PCA are 
akin to saliences in PLSC. Bootstrap and permutation tests are added to the 
analysis when inferential PSLC is indicated. 

A clean start is achieved by removing data and values from the working 
environment, closing all open graphic devices, and reporting the memory usage. 
The latter function, garbage collection, may help R return memory to the 
operating system after removing a large object. 

```{r clean start 8, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE}

rm(list = ls())
graphics.off()
gc()
```

```{r libraries for PLSC, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE}
library(ExPosition)
library(TExPosition)
#devtools::install_github('HerveAbdi/data4PCCAR') # if needed
library(data4PCCAR)
library(PTCA4CATA)
#install.packages('TInPosition') # if needed
library(TInPosition)
library(tidyverse)
library(corrplot)
library(prettyGraphs)
library(knitr)
library(kableExtra)

```

The following code refines the ggplot2 function by removing error messages
for multiple overlaps. 

```{r ggrepel, echo=FALSE, warning=FALSE,message=FALSE}
options(ggrepel.max.overlaps = Inf)
```


```{r, echo=FALSE, warning=FALSE,message=FALSE}
#Loading the data
hobbies = read.csv("data/hobbies.csv")

#rearrange row names outside the data matrix
hobbies_2 = hobbies[,-1]
rownames(hobbies_2) = hobbies[,1]

#omit observations with missing data
hobbies_3 = na.omit(hobbies_2)

#shorten variable names to make graphics easier to read
colnames(hobbies_3)=c("Hist","Psych","Politics","Math","Physics","Internet",  
                      "PC", "Econ","Bio","Chem","Read","Geo","Lang","Med","Law",
                      "Cars","Art","Religion", "Outdoor","Dance","Music",
                      "Writing","Passive Sp","Actice Sp","Garden","Celeb",
                      "Shop","STEM","Theatre","Friends","Adrenaline Sp", 
                      "Pets","Gender")

#I have roughly 100 more females than males in my data. I commented out the 
#code since it play little role in the analysis 

#hobbies_3[,'Gender'] <- plyr::mapvalues(hobbies_3[,'Gender'], 
                        #from = c("female", "male"), to = c("1","2")) 

#hobbies_3[,1:33] <- lapply(hobbies_3[,1:33], as.numeric)

#hist.gender = hist(hobbies_3[,33],breaks=20)
dim(hobbies_3)
```

```{r}
head(hobbies_3) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```

As a background to the data, I have a "hobbies" data set that is organized
into two natural sub-sets. The first thirteen were college major themes
(history, psychology, math) and the rest were post-college themes (law, medcine)
as well as lifelong hobbies (friends, sports, pets). Thus, I based my two 
matrices along a linear lifespan or life journey. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#splitting the data into two matrices

Xmat = hobbies_3[,1:13]

Ymat = hobbies_3[,14:32]

#grouping variable: gender

gender = hobbies_3[,33]
```

### Correlation plot

Politics and Law are highly correlated as is Physics and Science and 
Technology. PC and cars as well as PC and Science and Technology have
substantial correlations as well. Biology and Chemistry with Medicine have
very high correlations while Reading and Theater have a fair amount. An interest
in Reading is negatively correlated with an interest in Cars. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}

#correlation matrix

XY.cor = cor(Xmat, Ymat)

corrplot(XY.cor, method = "color", addCoef.col = "black", number.cex = .5)

 
```


### Analysis

```{r}
#Next I run the PLSC function

PLSC = tepPLS(Xmat, Ymat, DESIGN = gender, graphs = FALSE)

```

### Inferences

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#Iterations

nIter = 1000

#Permutation test
# Testing to see is results are non-random

Permutation.4.PLSC = perm4PLSC(
  Xmat,
  Ymat,
  nIter = 1000
)

#Bootstrap
#Testing to see is results are stable and would generate similar results
#in future analyses. 

Bootstrap.4.PLSC = Boot4PLSC(
  Xmat,
  Ymat,
  nIter = 1000,
  Fi = PLSC$TExPosition.Data$fi,
  Fj = PLSC$TExPosition.Data$fj,
  nf2keep = 3, #factors to keep for the analysis
  critical.value = 2, #cut-off value
  eig = TRUE, #compute CI for eigenvalues
  alphaLevel = .05 #alpha level for "eig"
  
)

```

### Graphics

```{r, echo=FALSE, warning=FALSE,message=FALSE}

#scree plot for eigenvalues
PlotScree(ev = PLSC$TExPosition.Data$eigs,
          title = "Hobbies and Gender: Inertia Scree Plot",
          plotKaiser = TRUE,
          color4Kaiser = ggplot2::alpha("blue",.5),
          lwd4Kaiser = 2)

#scree plot for singular values

PlotScree(ev = PLSC$TExPosition.Data$eigs^(1/2),
          title = "Hobbies and Gender: Singular Values Scree Plot",
          plotKaiser = FALSE)

```

### Color Scheme for Gender

```{r, echo=FALSE, warning=FALSE,message=FALSE}
Gcolors = gender

Gcolors = recode(Gcolors, "female" = "lightslategrey", 
                 "male" = "darkorchid1")
```

### Latent variable 1

```{r, echo=FALSE, warning=FALSE,message=FALSE}

laDim = 1

#combining latent variable x data with latent varibale y data

lv1.xy = cbind(
  PLSC$TExPosition.Data$lx[,laDim, drop = FALSE],
  PLSC$TExPosition.Data$ly[,laDim, drop = FALSE])

colnames(lv1.xy) = 
  c(paste0('LX',laDim), paste0('LY',laDim))

lv1 = createFactorMap(lv1.xy,
                      title = 'PLSC: First Pair of Latent Variables',
                      col.points = Gcolors,
                      alpha.points = .1,
                      col.labels = Gcolors,
                      alpha.labels = 0
                      )

LV1.a001 = lv1$zeMap+
  xlab(paste0("X Latent Variable ", laDim))+
  ylab(paste0("Y Latent Variable ", laDim))

# add group means to the factor map

Gmeans = PTCA4CATA::getMeans(lv1.xy, gender)

#recode to ensure colors match gender properly

col4Means = recode(rownames(Gmeans),
                  "female" = "lightslategrey", 
                 "male" = "darkorchid1")

#Factor map with means

MapMeans = PTCA4CATA::createFactorMap(Gmeans,
          constraints = lv1$constraints,
          col.points = col4Means,
          cex = 5, #size of the dot
          col.label = col4Means,
          text.cex = 4)

#Factor map with observations and group means

lv1.with.Means = LV1.a001+
  MapMeans$zeMap_dots + MapMeans$zeMap_text

print(lv1.with.Means)
```


### Confidence Intervals


```{r, echo=FALSE, warning=FALSE,message=FALSE}

#Bootstrap for the groups in the LV Space
# Bootstrap for CI

BootCube.Graph = PTCA4CATA::Boot4Mean(lv1.xy,
                design = gender,
                niter = 1000,
                suppressProgressBar = TRUE)

# The following creates the ellipses for the CI plots

dimnames(BootCube.Graph$BootCube)[[2]] = c("LX1","LY1")
GraphEllipse = PTCA4CATA::MakeCIEllipses(
  BootCube.Graph$BootCube[,1:2,],
  names.of.factors = c("LX1","LY1"),
  col = col4Means,
  p.level = .95
)

#latent variable map with observations, means, and confidence intervals

lv1.withCI = LV1.a001+MapMeans$zeMap_text+MapMeans$zeMap_dots+GraphEllipse
print(lv1.withCI)
```

### Contributions for LV1

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#Contribution for rows

#color scheme

Fi = PLSC$TExPosition.Data$fi

col4rows = prettyGraphsColorSelection(nrow(Fi),
                                      starting.color =1)
ctr.rows = PLSC$TExPosition.Data$ci

#contributions of the rows times the positive, neutral, or negative sign
# of the factor score of the rows. 
signed.ctri = ctr.rows * sign(Fi)

plot.Ctr.rows.1 = PrettyBarPlot2(
  bootratio = round(100*signed.ctri[,1]),
  threshold = 100/nrow(signed.ctri),
  ylim = c(-25,10),
  horizontal = FALSE,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4rows),
  color4ns = "gray75",
  plotnames = TRUE,
  main = "Important Contributions I-set: LV1",
  ylab = "Signed Contributions"
)
print(plot.Ctr.rows.1)

#same process for the columns

Fj = PLSC$TExPosition.Data$fj

col4columns = prettyGraphsColorSelection(nrow(Fj),
                                      starting.color =42)
ctr.columns = PLSC$TExPosition.Data$cj

signed.ctrj = ctr.columns * sign(Fj)

plot.Ctr.columns.1 = PrettyBarPlot2(
  bootratio = round(100*signed.ctrj[,1]),
  threshold = 100/nrow(signed.ctrj),
  ylim = c(-30,10),
  horizontal = FALSE,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4columns),
  color4ns = "gray75",
  plotnames = TRUE,
  main = "Important Contributions J-set: LV1",
  ylab = "Signed Contributions"
)
print(plot.Ctr.columns.1)
```

### Bootstrap

```{r, echo=FALSE, warning=FALSE,message=FALSE}
plotBR.rows.1 = PrettyBarPlot2(
  bootratio = Bootstrap.4.PLSC$bootRatios.i[,1],
  threshold = 2,
  ylim = NULL,
  horizontal = FALSE,
  signifOnly = TRUE, 
  color4bar = gplots::col2hex(col4rows),
  plotnames = TRUE,
  main = "Bootstrap Ratios. I-set: LV1",
  ylab = "Bootstrap Ratios"
)

print(plotBR.rows.1)

plotBR.columns.1 = PrettyBarPlot2(
  bootratio = Bootstrap.4.PLSC$bootRatios.j[,1],
  threshol = 2,
  ylim = NULL,
  horizontal = FALSE,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4columns),
  plotnames = TRUE,
  main = "Bootstrap Ratios. J-set: LV1",
  ylab = "Bootstrap Ratios"
)

print(plotBR.columns.1)
```

### Latent variable 2

```{r, echo=FALSE, warning=FALSE,message=FALSE}

laDim = 2

#combining latent variable x data with latent varibale y data

lv2.xy = cbind(
  PLSC$TExPosition.Data$lx[,laDim, drop = FALSE],
  PLSC$TExPosition.Data$ly[,laDim, drop = FALSE])

colnames(lv2.xy) = 
  c(paste0('LX',laDim), paste0('LY',laDim))

lv2 = createFactorMap(lv2.xy,
                      title = 'PLSC: Second Pair of Latent Variables',
                      col.points = Gcolors,
                      alpha.points = .1,
                      col.labels = Gcolors,
                      alpha.labels = 0
                      )

LV2.a001 = lv2$zeMap+
  xlab(paste0("X Latent Variable ", laDim))+
  ylab(paste0("Y Latent Variable ", laDim))

# add group means to the factor map

Gmeans.2 = PTCA4CATA::getMeans(lv2.xy, gender)

#recode to ensure colors match gender properly

col4Means.2 = recode(rownames(Gmeans.2),
                  "female" = "lightslategrey", 
                 "male" = "darkorchid1")

#Factor map with means

MapMeans.2 = PTCA4CATA::createFactorMap(Gmeans.2,
          constraints = lv2$constraints,
          col.points = col4Means.2,
          cex = 5, #size of the dot
          col.label = col4Means.2,
          text.cex = 4)

#Factor map with observations and group means

lv2.with.Means = LV2.a001+
  MapMeans.2$zeMap_dots + MapMeans.2$zeMap_text

print(lv2.with.Means)
```


### Confidence Intervals

```{r, echo=FALSE, warning=FALSE,message=FALSE}

#Bootstrap for the groups in the LV Space
# Bootstrap for CI

BootCube.Graph = PTCA4CATA::Boot4Mean(lv2.xy,
                design = gender,
                niter = 1000,
                suppressProgressBar = TRUE)

# The following creates the ellipses for the CI plots

dimnames(BootCube.Graph$BootCube)[[2]] = c("LX2","LY2")
GraphEllipse.2 = PTCA4CATA::MakeCIEllipses(
  BootCube.Graph$BootCube[,1:2,],
  names.of.factors = c("LX2","LY2"),
  col = col4Means.2,
  p.level = .95
)

#latent variable map with observations, means, and confidence intervals

lv2.withCI = LV2.a001+MapMeans.2$zeMap_text+MapMeans.2$zeMap_dots+GraphEllipse.2
print(lv2.withCI)
```

### Contributions for LV2

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#Contribution for rows

#color scheme

Fi.2 = PLSC$TExPosition.Data$fi

col4rows.2 = prettyGraphsColorSelection(nrow(Fi.2),
                                      starting.color =1)
ctr.rows.2 = PLSC$TExPosition.Data$ci

#contributions of the rows times the positive, neutral, or negative sign
# of the factor score of the rows.

signed.ctri.2 = ctr.rows.2 * sign(Fi.2)

plot.Ctr.rows.2 = PrettyBarPlot2(
  bootratio = round(100*signed.ctri.2[,2]),
  threshold = 100/nrow(signed.ctri.2),
  ylim = c(-25,10),
  horizontal = FALSE,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4rows.2),
  plotnames = TRUE,
  main = "Important Contributions I-set: LV2",
  ylab = "Signed Contributions"
)
print(plot.Ctr.rows.2)

#same process for the columns

Fj.2 = PLSC$TExPosition.Data$fj

col4columns.2 = prettyGraphsColorSelection(nrow(Fj.2),
                                      starting.color =42)
ctr.columns.2 = PLSC$TExPosition.Data$cj

signed.ctrj.2 = ctr.columns.2 * sign(Fj.2)

plot.Ctr.columns.2 = PrettyBarPlot2(
  bootratio = round(100*signed.ctrj.2[,2]),
  threshold = 100/nrow(signed.ctrj.2),
  ylim = c(-30,10),
  horizontal = FALSE,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4columns.2),
  plotnames = TRUE,
  main = "Important Contributions J-set: LV2",
  ylab = "Signed Contributions"
)
print(plot.Ctr.columns.2)
```

### Bootstrap

```{r, echo=FALSE, warning=FALSE,message=FALSE}
plotBR.rows.2 = PrettyBarPlot2(
  bootratio = Bootstrap.4.PLSC$bootRatios.i[,2],
  threshold = 2,
  ylim = NULL,
  horizontal = FALSE,
  signifOnly = TRUE, 
  color4bar = gplots::col2hex(col4rows),
  plotnames = TRUE,
  main = "Bootstrap Ratios. I-set: LV2",
  ylab = "Bootstrap Ratios"
)

print(plotBR.rows.2)

plotBR.columns.2 = PrettyBarPlot2(
  bootratio = Bootstrap.4.PLSC$bootRatios.j[,2],
  threshol = 2,
  ylim = NULL,
  horizontal = FALSE,
  signifOnly = TRUE,
  color4bar = gplots::col2hex(col4columns),
  plotnames = TRUE,
  main = "Bootstrap Ratios. J-set: LV2",
  ylab = "Bootstrap Ratios"
)

print(plotBR.columns.2)
```



### Conclusion 

Biology and Chemistry are highly correlated with Medicine while Reading is 
negatively correlated with Cars.Males and Females form distinct groups 
along the X and Y LV in both dimensions.Biology, Chemistry, Medicine correspond 
to the female group on LV1.Physics, Math, and Politics correspond to the male 
group on LV2 as does Adrenaline Sport, Cars, and Law. Shopping is a quality of 
females on LV2.



