
# DISTATIS {#distatis}

This method combines Multi-Dimensional Scaling and STATIS. The STATIS step follows
the MDS step and is an optimization step. Thus, optimum weights are added to the 
data table. Further, the sqaure of Eucledian distances are used to group variable
for each matrix. DiSTATIS is specifically a I x I x K matrix where I are objects
and K are people. 

```{r, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE}
rm(list = ls())
graphics.off()
gc()
```

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#devtools::install_github('HerveAbdi/PTCA4CATA')
#devtools::install_github('HerveAbdi/R4SPISE2018')
#devtools::install_github("HerveAbdi/R4SPISE2018" 
        #, dependencies = TRUE, 
    # use it first time only, comment after
    #build_opts = c("--no-resave-data", 
    # "--no-manual"), # Explicitly get the option
    #build_vignettes = TRUE,
    #force = TRUE)
#devtools::install_github('HerveAbdi/DistatisR')
#devtools::install_github('HerveAbdi/data4PCCAR')
#install.packages("prettyGraphs")
#install.packages('Matrix')
#install.packages('factoextra')
#install.packages('ExPosition')
#install.packages("DistatisR")
suppressMessages(library(Matrix))
suppressMessages(library(factoextra))
suppressMessages(library(RColorBrewer))
suppressMessages(library(DistatisR))
suppressMessages(library(PTCA4CATA))
suppressMessages(library(prettyGraphs))
suppressMessages(library(ExPosition))
library(R4SPISE2018)
suppressMessages(library(data4PCCAR))
library(knitr)
library(kableExtra)
```

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#Loading the data
wine = read.csv("Data/wine.csv")

data_2=wine[,-1]
rownames(data_2)=wine[,1]

data_3 = data.frame(data_2)

data_4 = t(data_3)

data_5 = data_3[-15,]


nVarJudges  <- ncol(data_4)

k <- 15

descJudges <- data_4[,k ]
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
## ----colJudges----------------------------------
# Create a 0/1 group matrix 
# with ExPosition::makeNominalData()
nominal.Judges <- makeNominalData(
                       as.data.frame(descJudges))
# get the colors
color4Judges.list <- 
    prettyGraphs::createColorVectorsByDesign(
                  nominal.Judges)

# color4Judges.list
color4Judges.list$oc <- descJudges

col4FI <- "red"
col4FnoI = "pink"
col4SAI = "black"
col4SAnoI<- 'blue'
color4Judges.list$oc[
       color4Judges.list$oc == 'Fr Info'] <- col4FI
color4Judges.list$oc[
       color4Judges.list$oc == 'Fr No Info'] <- col4FnoI
color4Judges.list$oc[
       color4Judges.list$oc == 'SA Info'] <- col4SAI
color4Judges.list$oc[
       color4Judges.list$oc == 'SA No Info'] <- col4SAnoI
color4Judges.list$gc[1,1] <- col4FI
color4Judges.list$gc[2,1] <- col4FnoI
color4Judges.list$gc[3,1] <- col4SAI
color4Judges.list$gc[4,1] <- col4SAnoI


DistanceCube <- DistanceFromSort(data_5)
```

```{r}
head(DistanceCube) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```

## Analysis 


```{r}
# **** Computations ----
## runDistatis--------------------------------
resDistatis <- distatis(DistanceCube, 
                        nfact2keep = 10)

n.active <- dim(DistanceCube)[3]

```

```{r, echo=FALSE, warning=FALSE,message=FALSE}
## rvGroups-----------------------------------
# Get the factors from the Cmat analysis
G <- resDistatis$res4Cmat$G 

# Compute the mean by groups of Judges
JudgesMeans.tmp <- aggregate(G, 
                      list(descJudges), mean) 
JudgesMeans <- JudgesMeans.tmp[,2:ncol(
                             JudgesMeans.tmp )] 
rownames(JudgesMeans) <- JudgesMeans.tmp[,1]
# Get the bootstrap estimates
BootCube <- PTCA4CATA::Boot4Mean(G, 
                       design = descJudges,
                       niter = 100,
                       suppressProgressBar = TRUE)
# head(BootCube)
# ## ----computeSk/ Fk ---------------------------
# Use Distatis1.1.0 function computePartial4Groups
groupK <- computePartial4Groups(
   resDistatis = resDistatis,
   DESIGN = descJudges
)
alpha_k <- groupK $groupAlpha
F_k     <- groupK $groupFS
```

## Graphs

The Rv coefficient is a correlation coefficient for covariance matrices.

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#**** Graphs ----
# ...  Rv Graphs ----
## Rv Corplot ----
# Note that the assessors have been re-ordered
# according to their projection on Dimension 1.
plot <- corrplot::corrplot(resDistatis$res4Cmat$C, 
                  # title = "The RV Map",
   order = 'FPC', # order from the first PC
   method = "color",
   col = RColorBrewer::brewer.pal(n = 8, 
                                  name = "PuOr"),
   number.cex = 0.8, 
   tl.col = color4Judges.list$oc,
   mar = c(0,0,0,0),
   addgrid.col = "grey", 
   tl.srt = 50)
```


The Rv Graph shows that French Judges with Information are likely to judge
similarly to other French Judges with Information as are South African 
judges and other South African judges. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}

## ----RV.scree.MapPlain -------
# 5.A. A scree plot for the RV coef. 
# Using standard plot (PTCA4CATA)
scree.rv.out <- PlotScree(
          ev = resDistatis$res4Cmat$eigValues, 
          title = "RV-map: Scree")

## ----RVGplot------------------------------------
# Create the layers of the map
RV.constraints <- minmaxHelper(
               resDistatis$res4Cmat$G[,1:2])
gg.rv.graph.out <- createFactorMap(
             X = resDistatis$res4Cmat$G, 
             axis1 = 1, axis2 = 2, 
             title = "Judges: RVMap", 
             col.points = color4Judges.list$oc, 
             col.labels = color4Judges.list$oc,
             constraints = RV.constraints)
# create the labels 
# for the dimensions of the RV map
labels4RV <- createxyLabels.gen(
        lambda = resDistatis$res4Cmat$eigValues, 
        tau    = resDistatis$res4Cmat$tau,
        axisName = "Dimension ")
## Create the map from the layers
# Here with labels and dots
a2a.gg.RVmap <- gg.rv.graph.out$zeMap + labels4RV
# Here with colored dots only
a2b.gg.RVmap <- gg.rv.graph.out$zeMap_background +
                gg.rv.graph.out$zeMap_dots + 
                labels4RV
a2a.gg.RVmap 
a2b.gg.RVmap
```

Judges, surprisingly, do not separate along the second dimension with respect
to nationality. 

```{r}
## ----RVwithCI-----------------------------------
# First the means
# A tweak for colors
in.tmp    <- sort(rownames(color4Judges.list$gc), 
                        index.return = TRUE)$ix
col4Group <- color4Judges.list$gc[in.tmp]
#
gg.rv.means <- PTCA4CATA::createFactorMap(
        JudgesMeans,
        axis1 = 1, axis2 = 2, 
        constraints = gg.rv.graph.out$constraints,
        col.points =  col4Group ,
        alpha.points = 1, # no transparency
        col.labels = col4Group)
#
 dimnames(BootCube$BootCube)[[2]] <- 
     paste0('dim ',1: dim(BootCube$BootCube)[[2]])
  #c('Dim1','Dim2') 
GraphElli.rv <- MakeCIEllipses(
        BootCube$BootCube[,1:2,],
        names.of.factors = c("dim 1","dim 2"), 
        col = col4Group, 
        p.level = .95)
a2d.gg.RVMap.CI <- a2b.gg.RVmap +
                  gg.rv.means$zeMap_dots + 
                  GraphElli.rv 
# dev.new()
print(a2d.gg.RVMap.CI)
## ----meansRV------------------------------------
#knitr::kable(JudgesMeans[,1:3])
## ----mapa2d -----------------
#print(a2d.gg.RVMap.CI )
```

The confidence intervals overlap so groups are not distinct. 


```{r}
##  HCA
##   Hierarchical Clustering Analysis (HCA)
 D <- dist(resDistatis$res4Cmat$G,
               method = "euclidean")
 fit <- hclust(D, method = "ward.D2")
 a05.tree4participants <- fviz_dend(fit,  
     k = 1, 
     k_colors = 'burlywood4', 
     label_cols = color4Judges.list$oc[fit$order],
     cex = .7, xlab = 'Participants',
     main = 'Cluster Analysis: Participants') 

## ---- plothca --------------------
print(a05.tree4participants)
```


No apparent pattern emerges from the Hierarchnical Clustering Analysis. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
 # ... S graphs ----
# S corplot ----
 #-________________________________________________
## CorrMat 4  Products  ----
 # Create color for the Products from prettyGraph
 color4Products <- #  
   prettyGraphsColorSelection(n.colors = 
                nrow(resDistatis$res4Splus$F))
 plot <- corrplot::corrplot(
   resDistatis$res4Splus$Splus, 
   #title = "The S Map",
   order = 'original', # original order
   method = "color",
   col = brewer.pal(n = 8, name = "PRGn"),
   number.cex = 0.8, 
   tl.col = color4Products,
   mar = c(0,0,0,0),
   addgrid.col = "grey", 
   tl.srt = 50)

# Reordered from HCA
 plot <- corrplot::corrplot(
   resDistatis$res4Splus$Splus, 
   #title = "The S Map",
   order = 'hclust', # order from a HCA
   method = "color",
   col = brewer.pal(n = 8, name = "PRGn"),
   number.cex = 0.8, 
   tl.col = color4Products,
   mar = c(0,0,0,0),
   addgrid.col = "grey", 
   tl.srt = 50)
```

FBAU and FCLL form a cluster of similarity as do SBEA, FCAR, and FROC as well as
FCLP and FROU. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
 ## ---- scree4S --------------------
 #_________________________________________________
 # A scree plot for the Compromise.
 scree.S.out <- PlotScree(
   ev = resDistatis$res4Splus$eigValues, 
   title = "Compromise: Explained Variance per Dimension")
```


```{r, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE}
## ----createGr4S -----
# 4.1 Get the bootstrap factor scores 
 #    (with default 1000 iterations)
BootF <- BootFactorScores(
                resDistatis$res4Splus$PartialF)

```


```{r, echo=FALSE, warning=FALSE,message=FALSE}
# 5.2 a compromise plot
# General title for the compromise factor plots:
genTitle4Compromise = 'Compromise.'
# To get graphs with axes 1 and 2:
h_axis = 1
v_axis = 2
# To get graphs with say 2 and 3 
# change the values of v_axis and h_axis
gg.compromise.graph.out <- createFactorMap(
                resDistatis$res4Splus$F,
                axis1 = h_axis, 
                axis2 = v_axis,
                title = genTitle4Compromise,
                col.points = color4Products ,
                col.labels = color4Products) 
# NB for the lines below 
#  You need DISTATIS version > 1.0.0
#  to get the eigen values 
#   and tau for the compromise
label4S <- createxyLabels.gen(
      x_axis   = h_axis, y_axis = v_axis,
      lambda   = resDistatis$res4Splus$eigValues, 
      tau      = resDistatis$res4Splus$tau,
      axisName = "Dimension ")
b2.gg.Smap <-  gg.compromise.graph.out$zeMap + 
                   label4S 
# dev.new()
print(b2.gg.Smap)
```


FROC, FCAR, SBEA, and SKLE form a distinct cluster along dimension 1 while
SRAA, FBAU, and SFED form a distinct cluster slong dimension 2. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#  
# 5.4 a bootstrap confidence interval plot 
# 5.3  create the ellipses
## With Ellipses -----
gg.boot.graph.out.elli <- MakeCIEllipses(
           data = BootF[,c(h_axis,v_axis),],
           names.of.factors = 
                c(paste0('Factor ',h_axis),
                  paste0('Factor ',v_axis)),
           col = color4Products,
)  
# Add ellipses to compromise graph
b3.gg.map.elli <- gg.compromise.graph.out$zeMap +
                  gg.boot.graph.out.elli + label4S 
#
## ----plot with ellipse -----------
print(b3.gg.map.elli)
```


```{r}
##  HCA products ------
nFac4Prod = 5
D4Prod <- dist(
           resDistatis$res4Splus$F[,1:nFac4Prod],
           method = "euclidean")
 fit4Prod <- hclust(D4Prod, method = "ward.D2")
 b3.tree4Product <- fviz_dend(fit4Prod,  k = 1, 
       k_colors = 'burlywood4', 
      label_cols = color4Products[fit4Prod$order],
       cex = .7, xlab = 'Products',
       main = 'Cluster Analysis: Products') 

## ----plot hca Prod ---------------------
 print(b3.tree4Product)
```

There seems like four major clusters with the first (SRAA, FBAU, and SFED)
and fourth (FROC, FCAR, SBEA, and SKLE) corresponding well with the projections 
along dimensions 1 and 2 seen earlier. 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
## ----PartialFS----------------------------------
# get the partial map

map4PFS <- createPartialFactorScoresMap(
         factorScores = resDistatis$res4Splus$F,      
         partialFactorScores = F_k,  
         axis1 = 1, axis2 = 2,
         colors4Blocks = color4Judges.list$gc, 
         colors4Items = as.vector(color4Products), 
         names4Partial = dimnames(F_k)[[3]], # 
         font.labels = 'bold')
# partial maps
d1.partialFS.map.byProducts <- 
          gg.compromise.graph.out$zeMap + 
          map4PFS$mapColByItems + label4S 
d2.partialFS.map.byCategories  <- 
          gg.compromise.graph.out$zeMap + 
          map4PFS$mapColByBlocks + label4S 

## ----S with Categories -----
print(d1.partialFS.map.byProducts )
## ----S with Categories.2 -----
print(d2.partialFS.map.byCategories)
```

### Summary

Two central cluster were seen in the analysis with the first (SRAA, FBAU, and SFED)
being seen along the first dimension and the fourth (FROC, FCAR, SBEA, and SKLE) 
be seen along the second dimension. 


