
# MFA {#mfa}

This method is an extension of PCA with a multi-data table scenario. First, MFA
does a PCA on each Table and normalizes each one. Second, the normalize tables
are aggregated in a multi-dimensional table and another non-normalized PCA is 
run to generate factor scores and loadings. MFA specifically is a I by J by K matrix
where I and J are objects and K are people. 

```{r start, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE}


# Clean Start ---- 
rm(list = ls())
graphics.off()
gc()
```

```{r libraries, echo=FALSE, warning=FALSE,message=FALSE}
#
#_________________________________________________
#  *** Prelude ----
# Libraries ----
#install.packages("prettyGraphs")
#install.packages("survival")
library(survival)
library(usethis)
library(devtools)
library(prettyGraphs)
library(ExPosition)
library(InPosition)
library(TExPosition)
library(TInPosition)
library(colortools)
library(coin)
library(abind)
library(carData)
library(car)
library(stats)
library(gplots)
library(DistatisR)
library(FactoMineR)
library(PTCA4CATA)
library(data4PCCAR)
library(knitr)
library(kableExtra)

```


```{r data, echo=FALSE, warning=FALSE,message=FALSE}
#Loading the data
wine = read.csv("Data/wine2.csv")

data_2=wine[,-1]
rownames(data_2)=wine[,1]

data_3 = data.frame(data_2)

ratings = data_3

nationVector = c("Fr.Info", "Fr.Info", "Fr.Info", "Fr.Info", "Fr.Info", "Fr.Info", 
"Fr.Info", "Fr.Info", "Fr.Info", "Fr.Info", "Fr.Info", "Fr.Info", "Fr.No.Info",
"Fr.No.Info","Fr.No.Info","Fr.No.Info","Fr.No.Info","Fr.No.Info","Fr.No.Info",
"Fr.No.Info","Fr.No.Info","Fr.No.Info","Fr.No.Info","Fr.No.Info","Fr.No.Info",
"SA.Info","SA.Info","SA.Info","SA.Info","SA.Info","SA.Info","SA.Info","SA.Info",
"SA.Info","SA.Info","SA.Info","SA.Info","SA.Info","SA.No.Info","SA.No.Info",
"SA.No.Info","SA.No.Info","SA.No.Info","SA.No.Info","SA.No.Info","SA.No.Info")
nVar4Nation = c("Fr.Info" = 12, "Fr.No.Info"  = 13, "SA.Info"=13, 
                "SA.No.Info" = 8)
namesOfNations = c("Fr.Info", "Fr.No.Info" , "SA.Info", 
                "SA.No.Info")
color4Wines = c("mediumorchid4", "olivedrab3", "palegreen1", "plum", 
                "royalblue", "mediumorchid4", "olivedrab3", "palegreen1", 
                "plum", "royalblue", "mediumorchid4", "olivedrab3", 
                "palegreen1", "plum")
color4Nations = c("deeppink","firebrick4", "honeydew4", "lemonchiffon3")

color4WinesB = as.matrix(color4Wines)
color4NationsB = as.matrix(color4Nations)
my_list = list(Ratings = ratings, NationVector = nationVector, 
               nVar4Nation2 = nVar4Nation, NamesOfNations = namesOfNations,
               Color4WinesB = color4WinesB, Color4NationsB = color4NationsB)

```

```{r, echo=FALSE, warning=FALSE,message=FALSE}

# Run MFA ----
# get the parameters
nNations5     <- length(my_list$NamesOfNations)
nProducts5   <- nrow(my_list$Ratings)
nVar4Nations5 <- my_list$nVar4Nation2
ratings5     <- my_list$Ratings
namesOfNations5 <- my_list$NamesOfNations
```

```{r}
head(ratings5) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)

```


## Analysis 


```{r, warning=FALSE, message=FALSE}
## call  MFA ----
resMFA <- FactoMineR::MFA(ratings5,
     group = nVar4Nations5,
     type = rep("s", nNations5),
     name.group = my_list$NamesOfNations,
     graph = FALSE  # TRUE first pass only
)


```

```{r, include = FALSE, echo=FALSE, warning=FALSE,message=FALSE}

# Run Distatis ----
## Distances
# Get the cube of distances for distatis
cubeOfDistances <- array(data = NA,
      dim = c(nProducts5, nProducts5, nNations5),
      dimnames = list(rownames(ratings5), 
      rownames(ratings5), namesOfNations5)
)

lindex <- cumsum(c(0,nVar4Nations5))

for (k in 1: nNations5){
  mat_k <- as.matrix(ratings5[,(lindex[k] + 1) : lindex[k + 1]])
  print(k)
  print(mat_k)
  norm_mat_k <- scale0(mat_k) 
  # NB Squared Euclidean
  cubeOfDistances[,,k] <- as.matrix(dist(norm_mat_k)^2)
}
```

```{r}

# distatis ----
resDistatis <- distatis(cubeOfDistances)
# **** Graphs ----
color4NationsB   <- my_list$Color4NationsB
color4WinesB<- my_list$Color4WinesB
# ... MFA ----
val.p   <- resMFA$eig[,1]
val.tau <- resMFA$eig[,2] 
ctr.Judges.mfa <- resMFA$group$coord

```

### Graphs 

```{r, echo=FALSE, warning=FALSE,message=FALSE}
# Scree ----
#_________________________________________________
scree.mfa <- PlotScree(
  ev = val.p, 
  title = "MFA: Explained Variance per Dimension")

#_________________________________________________
## ----RVGplot------------------------------------
# get the eigenvalues for RV
RV.eig <- eigen(resMFA$group$RV[1:nNations5, 1:nNations5], 
                symmetric = TRUE)
G.mfa  <- firstpos(RV.eig$vectors) %*% 
                        diag(RV.eig$values^(1/2))
rownames(G.mfa) <- namesOfNations5
colnames(G.mfa) <- paste0('Dimension ', 1:ncol(G.mfa))
mfa.rv.eig <- RV.eig$values
mfa.rv.tau <-  round(100 * mfa.rv.eig / sum(mfa.rv.eig))
# Create the layers of the map
gg.rv.graph.out.mfa <- createFactorMap(
  X = as.data.frame(G.mfa), 
  axis1 = 1, axis2 = 2, 
  title = "MFA. Judges: RVMap", 
  col.points = color4NationsB, 
  col.labels = color4NationsB)
# create the labels for the dimensions of the RV map
labels4RV.mfa <- createxyLabels.gen(
  lambda =  mfa.rv.eig, 
  tau    =  mfa.rv.tau,
  axisName = "Dimension ")
# # Create the map from the layers
# Here with labels and dots
a.a2a.gg.RVmap.mfa <- gg.rv.graph.out.mfa$zeMap + 
                        labels4RV.mfa
print(a.a2a.gg.RVmap.mfa)

```

Interestingly, Nationality and Amount of Information are not grouped together. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
#_________________________________________________
# Global Factor Scores ----
constraints.mfa <- minmaxHelper(resMFA$ind$coord.partiel)
F.mfa <- resMFA$ind$coord
# To get graphs with axes 1 and 2:
h_axis = 1
v_axis = 2
genTitle4Compromise = 'Compromise / Global Map. mfa'
gg.compromise.graph.out.mfa <- createFactorMap(
  F.mfa,
  axis1 = h_axis, 
  axis2 = v_axis,
  title = genTitle4Compromise,
  col.points = color4WinesB ,
  col.labels = color4WinesB,
  constraints = constraints.mfa)
label4S.mfa <- createxyLabels.gen(
  x_axis   = h_axis, y_axis = v_axis,
  lambda   = resMFA$eig[,1] , 
  tau      = round(resMFA$eig[,2]),
  axisName = "Dimension ")
b2.gg.Smap.mfa <-  
  gg.compromise.graph.out.mfa$zeMap + label4S.mfa 
print(b2.gg.Smap.mfa)

```

SKLE, FCLL, FCLP, and FROU are separated from SRAD, and SREY along the first dimension. 

SRAA, SFED, and FBAU are separated from FCAR, and FHUE along the second dimension. 


```{r, echo=FALSE, warning=FALSE,message=FALSE}
#________________________________________________
# Partial FS ----
# reformat ----
# Format MFA results to match distatis
# Partial coordinates in MFA
F_long <- resMFA$ind$coord.partiel 
F_k.mfa <-  array(data = NA,
                  dim = c(nProducts5, ncol(F_long), nNations5),
                  dimnames = list(rownames(ratings5), 
                                  colnames(F_long), namesOfNations5)
)
for (k in 1 : nNations5){
  row2keep <- seq(k, nProducts5*nNations5, nNations5)
  F_k.mfa[,,k] <- F_long[row2keep,]
}
map4PFS.mfa <- createPartialFactorScoresMap(
  factorScores = F.mfa,      
  partialFactorScores = F_k.mfa,  
  axis1 = 1, axis2 = 2,
  colors4Items = as.vector(color4WinesB), 
  names4Partial = dimnames(F_k.mfa)[[3]], # 
  font.labels = 'bold'
)
d1.partialFS.map.mfa.byProducts <- 
  gg.compromise.graph.out.mfa$zeMap + 
  map4PFS.mfa$mapColByItems + label4S.mfa 
d2.partialFS.map.mfa.byCategories  <- 
  gg.compromise.graph.out.mfa$zeMap + 
  map4PFS.mfa$mapColByBlocks + label4S.mfa 
print(d1.partialFS.map.mfa.byProducts)
print(d2.partialFS.map.mfa.byCategories)
```

```{r, echo=FALSE, warning=FALSE,message=FALSE}
#_________________________________________________
# Here we show the Circle of Correlation
# All J Circle ----
#_________________________________________________
# Compute correlation between variables & factors
cor.ratings <- cor(ratings5, F.mfa)

col4J  <- rep(color4NationsB, times = nVar4Nations5)
jolie.ggplot.J <- PTCA4CATA::createFactorMap(
  cor.ratings,
  col.points = col4J, col.labels = col4J, 
  constraints = list(minx = -1, miny = -1,
                     maxx = 1 , maxy = 1)   )
# draw the circle
e1.jolieggMap.J <- jolie.ggplot.J$zeMap + 
  addCircleOfCor() + label4S.mfa
print(e1.jolieggMap.J) 
#  Add some arrows
arrows <- addArrows(cor.ratings, color = col4J)  
e2.jolieggMap.J <- e1.jolieggMap.J + 
  arrows 
print(e2.jolieggMap.J)
# no dots
e3.jolieggMap.J <- jolie.ggplot.J$zeMap_background +
  jolie.ggplot.J$zeMap_text + arrows +
  addCircleOfCor() + label4S.mfa 
# Gray the small values ----
corLevels <- rowSums(cor.ratings[,1:2]^2) 
threshold <-  .50
col4J.gray <- col4J
col4J.gray[corLevels < threshold] <- 'gray85'
jolie.ggplot.J.gray <- PTCA4CATA::createFactorMap(
  cor.ratings,
  col.points = col4J.gray, col.labels = col4J.gray, 
  constraints = list(minx = -1, miny = -1,
                     maxx = 1 , maxy = 1)   )
arrows.gray <- addArrows(cor.ratings, 
                         color = col4J.gray)  
e8.jolieggMap.J.gray <- 
  jolie.ggplot.J$zeMap_background +
  jolie.ggplot.J.gray$zeMap_text + 
  arrows.gray +
  addCircleOfCor() + label4S.mfa 
print(e8.jolieggMap.J.gray)

```

```{r, echo=FALSE, warning=FALSE,message=FALSE}
# **** Your Turn: ----
# Do the graphs for Distatis and compare with MFA
# **** Graphs Distatis ----
# Distatis Graphs
# Scree ----
#_________________________________________________
scree.distatis <- PlotScree(
  ev = resDistatis$res4Cmat$eigValues, 
  title = "distatis: RV Explained Variance per Dimension")

#_________________________________________________
#_________________________________________________
scree.distatis.S <- PlotScree(
  ev = resDistatis$res4Splus$eigValues, 
  title = "distatis: Compromise Explained Variance per Dimension")

#_________________________________________________
## ----RVGplot------------------------------------
# get the eigenvalues for RV
G.distatis  <- firstpos(resDistatis$res4Cmat$G)
colnames(G.distatis) <- paste0('Dimension ',
                          1:ncol(G.distatis))
distatis.rv.eig <- resDistatis$res4Cmat$eigValues
distatis.rv.tau <-  round(100 *distatis.rv.eig / 
                          sum(distatis.rv.eig))
# Create the layers of the map
gg.rv.graph.out.distatis <- createFactorMap(
  X = as.data.frame(G.distatis), 
  axis1 = 1, axis2 = 2, 
  title = "DISTATIS. Judges: RVMap", 
  col.points = color4NationsB, 
  col.labels = color4NationsB)
# create the labels for the dimensions of the RV map
labels4RV.distatis <- createxyLabels.gen(
  lambda =  distatis.rv.eig, 
  tau    =  distatis.rv.tau,
  axisName = "Dimension ")
# # Create the map from the layers
# Here with labels and dots
f.a0002.a2a.gg.RVmap.distatis <- 
      gg.rv.graph.out.distatis$zeMap + 
      labels4RV.distatis
#_________________________________________________
# Global Factor Scores ----
constraints.distatis <- minmaxHelper(
  matrix(resDistatis$res4Splus$PartialF[,1:2,], 
         nrow = nProducts5 * nNations5, ncol = 2) )
F.distatis <- resDistatis$res4Splus$F
# To get graphs with axes 1 and 2:
h_axis = 1
v_axis = 2
genTitle4Compromise.distatis = 'Compromise / Global Map. DISTATIS'
gg.compromise.graph.out.distatis <- createFactorMap(
  F.distatis,
  axis1 = h_axis, 
  axis2 = v_axis,
  title = genTitle4Compromise.distatis,
  col.points = color4WinesB ,
  col.labels = color4WinesB,
  constraints = constraints.distatis)
label4S.distatis <- createxyLabels.gen(
  x_axis   = h_axis, y_axis = v_axis,
  lambda   = resMFA$eig[,1] , 
  tau      = round(resMFA$eig[,2]),
  axisName = "Dimension ")
f2.gg.Smap.distatis <-  
     gg.compromise.graph.out.distatis$zeMap + 
     label4S.distatis 
#________________________________________________
# Partial FS ----
# reformat ----
# Format MFA results to match distatis
# Partial coordinates in MFA

F_k.distatis <- resDistatis$res4Splus$PartialF
map4PFS.distatis <- createPartialFactorScoresMap(
  factorScores = F.distatis,      
  partialFactorScores = F_k.distatis,  
  axis1 = 1, axis2 = 2,
  colors4Items = as.vector(color4WinesB), 
  names4Partial = dimnames(F_k.distatis)[[3]], # 
  font.labels = 'bold'
)
h1.partialFS.map.distatis.byProducts <- 
  gg.compromise.graph.out.distatis$zeMap + 
  map4PFS.distatis$mapColByItems + 
  label4S.distatis 
h2.partialFS.map.distatis.byCategories  <- 
  gg.compromise.graph.out.distatis$zeMap + 
  map4PFS.distatis$mapColByBlocks + 
  label4S.distatis 
#_____________________________________________________________________
# Here we show the Circle of Correlation
# All J Circle ----
#_________________________________________________
# Compute correlation between variables & factors
cor.ratings.distatis <- cor(ratings, F.distatis)

col4J  <- rep(color4NationsB, times = nVar4Nations5)
jolie.ggplot.J.distatis <- PTCA4CATA::createFactorMap(
  cor.ratings.distatis,
  col.points = col4J, col.labels = col4J, 
  constraints = list(minx = -1, miny = -1,
                     maxx = 1 , maxy = 1)   )
# draw the circle
i1.jolieggMap.J.distatis <- jolie.ggplot.J.distatis$zeMap + 
  addCircleOfCor() + label4S.distatis
print(i1.jolieggMap.J.distatis) 
#  Add some arrows
arrows.distatis <- addArrows(cor.ratings.distatis, color = col4J)  
i2.jolieggMap.J.distatis <- i1.jolieggMap.J.distatis + 
  arrows.distatis 
print(i2.jolieggMap.J.distatis)
# no dots
i3.jolieggMap.J.distatis <- jolie.ggplot.J.distatis$zeMap_background +
  jolie.ggplot.J.distatis$zeMap_text + arrows.distatis +
  addCircleOfCor() + label4S.distatis 
# Gray the small values ----
corLevels <- rowSums(cor.ratings.distatis[,1:2]^2) 
threshold <-  .50
col4J.gray <- col4J
col4J.gray[corLevels < threshold] <- 'gray85'
jolie.ggplot.J.gray.distatis <- PTCA4CATA::createFactorMap(
  cor.ratings.distatis,
  col.points = col4J.gray, col.labels = col4J.gray, 
  constraints = list(minx = -1, miny = -1,
                     maxx = 1 , maxy = 1)   )
arrows.gray.distatis <- addArrows(cor.ratings.distatis, 
                                  color = col4J.gray)  
i8.jolieggMap.J.gray.distatis <- 
  jolie.ggplot.J.distatis$zeMap_background +
  jolie.ggplot.J.gray.distatis$zeMap_text + 
  arrows.gray.distatis +
  addCircleOfCor() + label4S.distatis 
print(i8.jolieggMap.J.gray.distatis)
```

Here, there is evidence that French and South African Judges that contribute 
more importantly to the analysis are separated by the first dimension. 


## Summary 

SKLE, FCLL, FCLP, and FROU correspond South African Judges and 
SRAD, and SREY correspond to French Judges. 
